---
title: "timeline"
output: html_document
---

# Changes

## 2022-12 
*remove cases over time
*trim the fat
*use color scheme for provs from cases

## 2022-06 changes
* Repeat previous plots with updated info
* Repeat color scheme
* Compare different intervention stringency measures


```{r}
library(gtools)
library(tidyverse)
library(tidyr)
library(data.table)
library(cowplot)
library(lubridate)
library(stringr)
library(ggplot2)
library(RColorBrewer)
library(ggplotify)
library(gridExtra)
library(grid)
library(lattice)
library(reshape2)

```


# PLOT CASES OVER TIME IN CANADA colored by prov, with map legend
## ins and outs
```{r}
#inputs 
# the day of most recent sample
data.date<-"2022-03-22" 
start.date<-"2020-11-01" #about when VOCs first started coming to Canada

#provincial case counts
prov.cases<-read.csv("Canada_cases/20220516_covid19_casesbyprov.csv") #most upddated

#color scheme 
globalPalette<-read.table("colors_out/global.colors.tsv",sep="\t")

#outputs
#Folder of images, tables
today<-Sys.Date()
f.out<-paste0(today,"_Results/")
if(!dir.exists(f.out)){dir.create(f.out)}

## make color scheme
glob.colz<-row.names(globalPalette)
globalPalette.ch<-as.character(globalPalette$globalPalette)
names(globalPalette.ch)<-glob.colz

#make scales
GlobColScale<-scale_colour_manual(name = "Location",values = globalPalette.ch,na.value="grey60")
GlobFillScale<-scale_fill_manual(name = "Location",values = globalPalette.ch,na.value="grey60")

#make one only for Can provs and Can itself
globalPalette.prov<-globalPalette.ch[1:14]
ProvColScale<-scale_colour_manual(name = "Location",values = globalPalette.prov,na.value="grey60")
ProvFillScale<-scale_fill_manual(name = "Location",values = globalPalette.prov,na.value="grey60")

#even smaller one for provs with seq data
globalPalette.prov2<-globalPalette.ch[c(1:9)] #could add Canada here
ProvColScale2<-scale_colour_manual(name = "Location",values = globalPalette.prov2,na.value="grey60")
ProvFillScale2<-scale_fill_manual(name = "Location",values = globalPalette.prov2,na.value="grey60")

## plotting theme
pubThemeDate<-theme(panel.grid.major.y = element_blank(), 
                    panel.grid.minor = element_blank(),
                    panel.background = element_rect("grey95"), 
                    axis.line = element_line(colour = "black"),
                    text=element_text(size=10,face="bold"),
                    legend.text=element_text(size=8),
                    axis.text.x=element_text(angle = 45,hjust=1,size=rel(1)))

scaleDate<-scale_x_date(date_breaks = "1 month", date_minor_breaks = "2 weeks", date_labels = "%b %Y",limits=as.Date(c(start.date,data.date)),expand=c(0,0))

xdates<-ymd(c("2020-12-01","2021-03-01","2021-06-01","2021-09-01","2021-12-01","2022-03-01"))
scaleDateBusy<-scale_x_date(breaks=xdates, date_labels = "%b %Y",
                            limits=as.Date(c(start.date,data.date)),
                            expand=c(0,0))

scaleDateFlex<-scale_x_date(date_labels = "%b %Y",expand=c(0,0))

## Variant colors
variant.colors.in<-"colors_out/variant_colors.tsv"
varPalette<-read.table(variant.colors.in,sep="\t")

## make color scheme
var.colz<-as.character(varPalette$vars.ch)
varPalette.ch<-row.names(varPalette)
names(varPalette.ch)<-var.colz
VOCColScale<-scale_colour_manual(name = "Variant",values = varPalette.ch,na.value=NULL)
VOCFillScale<-scale_fill_manual(name = "Variant",values = varPalette.ch,na.value=NULL)

### add separate groups for each omicron group
omis<-c("BA.1","BA.1.1","BA.2")
# https://www.colorhexa.com/ec9a91
omiPalette.ch<-c("#e47266","#e8867b","#f0aea7") 
names(omiPalette.ch)<-omis
# morevarPal<-c(varPalette.ch,omiPalette.ch)
# morevarPal<-morevarPal[-which(names(morevarPal)=="Omicron")]

VarColScale<-scale_colour_manual(name = "Variant",values = omiPalette.ch,na.value=NULL)
VarFillScale<-scale_fill_manual(name = "Variant",values = omiPalette.ch,na.value=NULL)


```


## plot of cases over time
```{r}
# prov.cases<- prov.cases  %>%
#   dplyr::filter(prname != "Repatriated travellers")
# head(prov.cases)
prov.cases$date<-as.Date(prov.cases$date)

#total sums by 2020-12-31
tot.df<-prov.cases[prov.cases$date==data.date,]
prov.ord<-tot.df[with(tot.df,order(tot.df$numtotal,decreasing = T)),c("prname")]
prov.ord<-prov.ord[-1] #canada
# prov.ord<-prov.ord[c(1:9,11,10,12:13,14)] #keep maritimes and territories together

#make a modified object
prov.cases2<-prov.cases
prov.cases2<-prov.cases2 %>% dplyr::filter(prname != "Canada")
prov.cases2$prname<-factor(prov.cases2$prname, levels=prov.ord)

#Make a composite that has individual 'numtoday' for observations before rolling average could be calculated
prov.cases2$avgtotal_mod<-prov.cases2$avgtotal_last7
for (i in 1:nrow(prov.cases2)){ 
  #if avg is NA or 0
  if(is.na(prov.cases2$avgtotal_mod[i]) | prov.cases2$avgtotal_mod[i]==0){
    #and if number cases today is not NA, use that
    if(!is.na(prov.cases2$numtoday[i])){
      prov.cases2$avgtotal_mod[i]<-prov.cases2$numtoday[i]
    }
  }
  }


p2<-prov.cases2 %>%
  ggplot()+
  geom_bar(aes(x=date,y=avgtotal_mod,group=prname,fill=prname),stat="identity",width=1)+
  labs(x=NULL, y="Average daily new diagnoses*")+
  pubThemeDate+
  theme(axis.title.y=element_text(size=rel(1.25),face="bold"),
        # legend.position=c(0.25,0.7),
        legend.position = "none",
        legend.text=element_text(size=rel(0.9)))+
  scaleDateBusy+
  ProvFillScale
  # guides(fill=guide_legend(title="Province",title.position = "top", 
  #                          keywidth = 1,keyheight = 1,ncol=2))+
  # annotate(geom="text",x=as.Date("2021-01-01"),y=8400,
  #            label="Second wave",size=2,fontface="italic",hjust=0,vjust=0,color="black")+
  # annotate(geom="text",x=as.Date("2021-03-30"),y=8600,
  #            label="Third wave",size=2,fontface="italic",hjust=0,vjust=0,color="black")+
  # annotate(geom="text",x=as.Date("2021-08-25"),y=4900,
  #            label="Fourth wave",size=2,fontface="italic",hjust=0,vjust=0,color="black")+
  # annotate(geom="text",x=as.Date("2021-12-05"),y=42300,
  #            label="Fifth wave",size=2,fontface="italic",hjust=0,vjust=0,color="black")
# p2

# print it with the Canada legend
ggdraw() +  
  draw_plot(p2) 
  # draw_plot(p.map,scale=0.32,halign = 0.25,valign=0.99)
  # draw_image("../../20210301_phylogeo/01_subSample/Results/PlainMap_GlobalPalette.png",
             # scale=0.71,halign = 0.46,valign=0.91)
ggsave(paste0(f.out,"AvgDailyDiagnosesByProv.png"),width=5,height=3)
```


# STRINGENCY
##ins and outs
```{r}
#read in stringency for oxford
ox<-read.csv("oxford-tracker/OxCGRT_latest.csv") #one row a day
# head(ox)
colnames(ox)
ox.can<-ox[which(ox$CountryName=="Canada"),c("Date","CountryName","RegionName","StringencyIndex")]
# table(ox.can$RegionName) #one per province and one unlabelled (all canada presume)
ox.can$RegionName[ which(ox.can$RegionName=="")]<-"Canada"

#convert dates
ox.can$Date<-as.Date(as.character(ox.can$Date),format="%Y%m%d")
# head(ox.can) 

y.lim<-c(15,95)
#PLOT ER UP
ggplot(ox.can)+
  geom_line(aes(x=Date,y=StringencyIndex,group=RegionName,color=RegionName))+
  ProvColScale+
  pubThemeDate+
 scaleDate+
   theme(legend.position = "right",
        legend.text=element_text(size=rel(0.6)),
        legend.title=element_text(size=rel(0.7)))+
  labs(x=NULL,y="Oxford Stringency Index")+
  scale_y_continuous(limits=y.lim, expand = c(0,0))+
  guides(color=guide_legend(title="Province or\nterritory",title.position = "top",
                           keywidth = 0.5,keyheight = 0.5,ncol=1))+
  scaleDateBusy

ggsave(paste0(f.out,"Oxford Stringency Index over time_AllProvs.png"),height=2.5,width=5,units = "in")

#only provs with data and order high to low stringency (which had the lowest)
maxstring<-ox.can %>%  
  dplyr::filter(Date <data.date) %>%
  dplyr::filter(Date >start.date) %>%
  dplyr::group_by(RegionName) %>% 
  dplyr::summarise(.groups="rowwise",
                  meanStringency=mean(StringencyIndex,na.rm=T),
                  minStringency=min(StringencyIndex,na.rm=T),
                  maxStringency=max(StringencyIndex,na.rm=T)) %>% as.data.frame()
maxstring<-maxstring[order(maxstring$meanStringency,decreasing = T),]
prov.ord.st<-maxstring$RegionName
ox.can$RegionName<-factor(ox.can$RegionName, levels=prov.ord.st)

prov.ord.st2<-prov.ord.st[which(prov.ord.st %in% names(globalPalette.prov2))]
ProvColScale3<-scale_colour_manual(name = "Location",
                                   values = globalPalette.prov2[match(prov.ord.st2,names(globalPalette.prov2))],
                                   na.value=NULL)

ox.can %>%
  dplyr::filter(RegionName %in% names(globalPalette.prov2)) %>%
  ggplot()+
  geom_line(aes(x=Date,y=StringencyIndex,group=RegionName,color=RegionName))+
  ProvColScale3+
  pubThemeDate+
   theme(legend.position = "right",
        legend.text=element_text(size=rel(0.6)),
        legend.title=element_text(size=rel(0.7)))+
  labs(x=NULL,y="Oxford Stringency Index")+
  scale_y_continuous(limits=y.lim, expand = c(0,0))+
  guides(color=guide_legend(title="Province",title.position = "top",
                           keywidth = 0.5,keyheight = 0.5,ncol=1))+
  scaleDateBusy

ggsave(paste0(f.out,"Oxford Stringency Index over time_SelectProvsCanada.png"),height=2.5,width=5,units = "in")

```

## Calculate deviation from mean by day
```{r}
#separate canada
ox.canada<-ox.can[ox.can$RegionName=="Canada",c(1,4)]
colnames(ox.canada)[2]<-paste0("Canada_",colnames(ox.canada)[2])

#make a restricted version of this
ox.can.dev<- ox.can %>%   
  dplyr::filter(Date <data.date) %>%
  dplyr::filter(Date >start.date) %>% 
  dplyr::filter(RegionName!="Canada")
ox.can.dev.mean<-ox.can.dev %>% dplyr::group_by(Date) %>% dplyr::summarize(meanStringency=mean(StringencyIndex)) %>%
  as.data.frame()

#join on canada by day
ox.can.dev<-ox.can.dev %>% dplyr::left_join(ox.canada, by="Date") %>%
  dplyr::left_join(ox.can.dev.mean, by="Date")

#calculate deviation from the Canadian stringency
#as a deviance or variance and as a % coeffiient of var (normalized to mean *100)
ox.can.dev<-ox.can.dev %>% 
  dplyr::mutate(dev_string=StringencyIndex-meanStringency) %>% #if positive then prov string higher, if negative than less stringent than Canada
  dplyr::mutate(coeff_var_string=(sqrt(abs(dev_string))/meanStringency)*100) 

#Plot the deviances and coeff of var over time
ox.can.dev %>%
  dplyr::filter(RegionName %in% names(globalPalette.prov2))%>%
  ggplot()+
  geom_line(aes(x=Date,y=dev_string,group=RegionName,color=RegionName))+
  ProvColScale3+
  pubThemeDate+
   theme(legend.position = "none",
        legend.text=element_text(size=rel(0.6)),
        legend.title=element_text(size=rel(0.7)))+
  labs(x=NULL,y="Stringency residuals")+
  scale_y_continuous(expand = c(0,0))+
  guides(color=guide_legend(title="Province",title.position = "top",
                           keywidth = 0.5,keyheight = 0.5,ncol=1))+
  scaleDateBusy
ggsave(paste0(f.out,"OxfordString_Residuals_SelectProvsCanada_noleg.png"),height=2.5,width=4,units = "in")

ox.can.dev %>%
  dplyr::filter(RegionName %in% names(globalPalette.prov2))%>%
  ggplot()+
  geom_line(aes(x=Date,y=coeff_var_string,group=RegionName,color=RegionName))+
  ProvColScale3+
  pubThemeDate+
   theme(legend.position = "right",
        legend.text=element_text(size=rel(0.6)),
        legend.title=element_text(size=rel(0.7)))+
  labs(x=NULL,y="%CV stringency")+
  scale_y_continuous(expand = c(0,0))+
  guides(color=guide_legend(title="Province",title.position = "top",
                           keywidth = 0.5,keyheight = 0.5,ncol=1))+
  scaleDateBusy
ggsave(paste0(f.out,"OxfordString_NormalDeviance_SelectProvsCanada.png"),height=2.5,width=5,units = "in")

#daily binary of above or below average
ox.can.dev<-ox.can.dev %>% 
  dplyr::mutate(binaryString=(dev_string<0))

## summary statistics 
ox.can.summary<-ox.can.dev %>%
  # dplyr::filter(RegionName %in% names(globalPalette.prov2))%>%
  dplyr::group_by(RegionName)%>%
  dplyr::summarise(meanString=mean(StringencyIndex,na.rm=T),
                   medianString=median(StringencyIndex,na.rm=T),
                   maxString=max(StringencyIndex,na.rm=T),
                   minString=min(StringencyIndex,na.rm=T),
                   #variation of variance
                   meanStringVar=mean(dev_string,na.rm=T),
                   medianStringVar=median(dev_string,na.rm=T),
                   maxStringVar=max(dev_string,na.rm=T),
                   minStringVar=min(dev_string,na.rm=T),
              
                   meanCV=mean(coeff_var_string,na.rm=T),
                   medianCV=median(coeff_var_string,na.rm=T),
                   maxCV=max(coeff_var_string,na.rm=T),
                   minCV=min(coeff_var_string,na.rm=T),
                   
                   #how often above or below mean
                   sumAbove=sum(which(binaryString==F)),
                   sumBelow=sum(which(binaryString==T)) ) %>%
  dplyr::mutate((percAbove=sumAbove/sum(sumAbove,sumBelow) )*100,
                  (percBelow=sumBelow/sum(sumAbove,sumBelow) )*100)

write.csv(ox.can.summary,paste0(f.out,"OxfordStringenceResidualSummary.csv"))
```

## Compare Oxford stringency to Bank of Canada data - more local considerations
```{r}
#read in 
boc<-read.csv("bank-of-canada/COVID-19_STRINGENCY_INDEX-form.csv")
#go from wide to long for all the regions
boc.long<-boc %>% pivot_longer(cols=2:ncol(boc))
colnames(boc.long)<-c("Date","RegionName","StringencyIndex")
boc.long$RegionName<-str_replace_all(boc.long$RegionName, "\\.", " ")

#merge with oxford
boc.long$method<-"boc"
ox.can$method<-"oxford"
boc.long$Date<-as.Date(boc.long$Date)
#remove these regions
boc.long<-boc.long %>%   
  dplyr::filter(!RegionName %in% c("Atlantic region ","Prairies region ")) 

all.string<-bind_rows(ox.can[,c(colnames(boc.long))],
                      boc.long)
head(all.string)

#plot together for provs
all.string %>% 
  # filter(RegionName=="Canada") %>%
  ggplot()+
  geom_line(aes(x=Date,y=StringencyIndex,group=method, color=method))+
  theme(legend.position = "right",
        legend.text=element_text(size=1))+
  pubThemeDate+
 scaleDate+
  labs(x=NULL,y="Stringency Index")+
  scale_y_continuous(limits=y.lim, expand = c(0,0))+
  facet_wrap(~RegionName,nrow=3)+
  guides(color=guide_legend(title="Index",title.position = "top",
                           keywidth = 0.5,keyheight = 1,ncol=1))+
  scaleDateBusy
ggsave(paste0(f.out,"StringencyIndex_BOCvsOX_Comparison.png"),width=8.5,height=5,units="in")


```

## Plot stringency by prov in one plot for the Bank of Canada data
```{r}
#order it by mean stringency
meanstring.boc<-boc.long %>%   
  dplyr::group_by(RegionName) %>% 
  dplyr::summarise(.groups="rowwise",
                   meanstring.bocency=mean(StringencyIndex,na.rm=T)) %>% as.data.frame()
meanstring.boc<-meanstring.boc[order(meanstring.boc$meanstring.bocency,decreasing = T),]
prov.ord.st<-meanstring.boc$RegionName
#order it nicely
# prov.ord2<-c("Canada",prov.ord)
# prov.ord2<-prov.ord2[-which(prov.ord2=="Repatriated travellers")]
boc.long$RegionName<-factor(boc.long$RegionName, levels=prov.ord.st)

ProvColorScale2<-scale_color_manual(name = "Location",
                                    values = rev( colorRampPalette(
                                      brewer.pal(name="YlOrRd",n=9))
                                      (length(prov.ord.st))))
max(boc.long$StringencyIndex,na.rm=T)
y.lim<-c(0,85)
#PLOT ER UP
boc.long %>%
  dplyr::filter(RegionName %in% names(globalPalette.prov2))%>%
  ggplot()+
  # #annotate the background
  # annotate(geom="rect",xmin = as.Date("2021-04-22"), xmax = as.Date("2021-09-26"),
  #          ymin = y.lim[1], ymax =y.lim[2], color="grey85",size=0,  fill="grey80")+
  # annotate(geom="text",x = as.Date("2021-05-30"),y=y.lim[2], vjust=1, hjust=0,
  #          label="Flight ban",fontface="italic",color="grey25",size=3)+
  geom_line(aes(x=Date,y=StringencyIndex,group=RegionName,color=RegionName))+
  theme(legend.position = "right",
        legend.text=element_text(size=1))+
  ProvColScale3+
  pubThemeDate+
 scaleDate+
  labs(x=NULL,y="Bank of Canada Stringency Index")+
  scale_y_continuous(limits=y.lim, expand = c(0,0))+
  guides(color=guide_legend(title="Province",title.position = "top",
                           keywidth = 0.5,keyheight = 1,ncol=1))+
  scaleDateBusy

ggsave(paste0(f.out,"BankofCan Stringency Index over time_ProvsCanada_2021.png"),height=3,width=6,units="in")
```

## Look at the IRPP data
```{r}
irpp<-read.csv("IRPP/Canada-COVID19-Strengency_IRPP.csv")
head(irpp)
irpp<-irpp[,c("Province.Territory","date","stringencyIndex")] #look at more columsn another time
colnames(irpp)<-c("RegionName","Date","StringencyIndex")
irpp$method<-"irpp"

#Merge onto the all string object
irpp$Date<-as.Date(irpp$Date)
all.string2<-bind_rows(all.string,irpp)

#Plot comparison again
#plot together for provs
y.lim<-c(0,95)

all.string2 %>% 
  # filter(RegionName=="Canada") %>%
  ggplot()+
  geom_line(aes(x=Date,y=StringencyIndex,group=method, color=method))+
  theme(legend.position = "right",
        legend.text=element_text(size=1),
        strip.text.x= element_text(size=rel(0.8))
        )+
  pubThemeDate+
 scaleDate+
  labs(x=NULL,y="Stringency Index")+
  scale_y_continuous(limits=y.lim, expand = c(0,0))+
  facet_wrap(~RegionName,nrow=3)+
  guides(color=guide_legend(title="Index",title.position = "top",
                           keywidth = 0.5,keyheight = 1,ncol=1))+
  scaleDateBusy
ggsave(paste0(f.out,"StringencyIndex_OXvsBOCvsIRPP_comparison.png"),width=8,height=5,units="in")

```

## Plot comparison across provs using IRPP
```{r}
#order it by mean stringency
meanstring.irpp<-irpp %>%   
  dplyr::group_by(RegionName) %>% 
  dplyr::summarise(.groups="rowwise",
                   meanstring.irppency=mean(StringencyIndex,na.rm=T)) %>% as.data.frame()
meanstring.irpp<-meanstring.irpp[order(meanstring.irpp$meanstring.irppency,decreasing = T),]
prov.ord.st<-meanstring.irpp$RegionName
#order it nicely
# prov.ord2<-c("Canada",prov.ord)
# prov.ord2<-prov.ord2[-which(prov.ord2=="Repatriated travellers")]
irpp$RegionName<-factor(irpp$RegionName, levels=prov.ord.st)

ProvColorScale2<-scale_color_manual(name = "Location",
                                    values = rev( colorRampPalette(
                                      brewer.pal(name="YlOrRd",n=9))
                                      (length(prov.ord.st))))
max(irpp$StringencyIndex,na.rm=T)
y.lim<-c(-5,95)
#PLOT ER UP
irpp %>%
  dplyr::filter(RegionName %in% names(globalPalette.prov2))%>%
  ggplot()+
  #annotate the background
  # annotate(geom="rect",xmin = as.Date("2021-04-22"), xmax = as.Date("2021-09-26"),
  #          ymin = y.lim[1], ymax =y.lim[2], color="grey85",size=0,  fill="grey80")+
  # annotate(geom="text",x = as.Date("2021-05-30"),y=y.lim[2], vjust=1, hjust=0,
  #          label="Flight ban",fontface="italic",color="grey25",size=3)+
  
  geom_line(aes(x=Date,y=StringencyIndex,group=RegionName,color=RegionName))+
  ProvColScale3+
  pubThemeDate+
 scaleDate+
   theme(legend.position = "right",
        legend.text=element_text(size=rel(0.6)),
        legend.title=element_text(size=rel(0.7)))+
  labs(x=NULL,y="IRPP Stringency Index")+
  scale_y_continuous(limits=y.lim, expand = c(0,0))+
  guides(color=guide_legend(title="Province",title.position = "top",
                           keywidth = 0.5,keyheight = 0.5,ncol=1))+
  scaleDateBusy

ggsave(paste0(f.out,"IRPP Stringency Index over time_ProvsCanada_2021_sml.png"),height=3,width=5,units = "in")
```

# OLD from 2020 analysis

## Look at change in string
```{r}
#make a column for daily change
ox.can$StringencyIndex.change<-NA
for (i in 1:nrow(ox.can)){
  if (i>1){
      ox.can$StringencyIndex.change[i]<-ox.can$StringencyIndex[i]-ox.can$StringencyIndex[i-1]
  }
}

#biggest change 
ox.can$Date [which(ox.can$StringencyIndex.change==max(ox.can$StringencyIndex.change,na.rm = T))] #"2020-03-17"
max(ox.can$StringencyIndex,na.rm = T) #90.74

#biggest change in 2021
ox.can2021<-ox.can[ox.can$Date > as.Date("2020-12-31"),]
ox.can2021$Date [which(ox.can2021$StringencyIndex.change==max(ox.can2021$StringencyIndex.change,na.rm = T))] #""2021-12-30"

max(ox.can$StringencyIndex,na.rm = T)
```

## Read in custom timeline 
```{r}
varevents<-read.csv(paste0("2023-02-02_Results/KeyVariantEvents2022.csv"))
# head(varevents)
varevents<-varevents[-which(varevents$Response.type=="Contextual event"),]
varevents$Date<-varevents$Start.date

positions<-seq(0,nrow(varevents), by=1)
directions <- c(1, 1)
line_pos <- data.frame(
    "position"=rep(positions, length.out=nrow(varevents)),
    "direction"=rep(directions, length.out=nrow(varevents)))

#sort var events byd ate
varevents<-varevents[with(varevents,order(varevents$Date)),]
#bind
can.3 <- bind_cols(varevents,line_pos)
can.3$Date<-as.Date(can.3$Date)
#better colors
# unique(varevents$Response.type) 
respcolz<-c("sienna","darkorange1","royalblue3")

p1<-ggplot()+
  labs(x=NULL,y=NULL)+
  #description and lines to axis
  geom_segment(data=can.3,aes(y=position,yend=-1, 
                              x=Date,xend=Date), #,color=Intervention.Category
               size=0.2,alpha=0.2)+
  geom_text(data=can.3,aes(x=Date,label=Description,y=position, color=Response.type), fontface="bold", size=2.4,hjust=0)+ 
  pubThemeDate+
  theme(axis.text.y=element_blank(),
        legend.position=c(0.1,0.8),
        axis.ticks.y=element_blank())+
  scale_x_date(breaks=xdates, date_labels = "%b %Y",
                            limits=as.Date(c(start.date,data.date)),
                            expand=expansion(add=c(0,370)))+  
  coord_cartesian(ylim=c(-1,nrow(varevents)))+
  scale_color_manual(values=respcolz)+
  scale_y_continuous(expand = expansion(add=c(0, 0)))+
  guides(color=guide_legend(title = "Federal\nCOVID-19\ninterventions"))

p1
ggsave(paste0(f.out,"Canada-intervention-timeline.png"),height=4.5,width=9.6,units="in")

```

## Make a simple version with only variant specific interventions
```{r}
can.var<-can.3  %>% dplyr::filter(Response.type=="Variant-specific")

can.var$Start.date<-as.Date(can.var$Start.date)
can.var$End.date<-as.Date(can.var$End.date)
can.var$Description[nrow(can.var)]<-  "Enhanced screening and quarantine for Canadians and banned\nentry of foreign nationals who had travelled to southern Africa"
#remove redundant
can.var2<-can.var[-(nrow(can.var)-1),]
can.var2$position<-1:nrow(can.var2)
can.var2$variant<-c("Alpha","Beta","Gamma","Gamma","Delta","Delta","Omicron")

#option to remove the ban against pakistan b/c not testing that
can.var2<-can.var2 %>%
  filter(End.date != as.Date("2021-06-08")) #filter out pakistan for summary

#save these
write.csv(can.var2,file=paste0(f.out,"can.variant.interv.plotty2.csv"))

#make y positions (same for two gamma interventions)
can.var2$position<-c(1,2,3,3,4,5)
# can.var2$position<-1:nrow(can.var2)

#remove redundant text for two gamma interves
can.var2$Description[3]<-NA
can.var2$variant2<-can.var2$variant #duplicate
can.var2$variant2[4]<-NA
# can.var2$variant[3]<-NA #need this text 
# head(can.var)
p2<-can.var2 %>%
  ggplot()+
  labs(x=NULL,y=NULL)+
  #description and lines to axis
  geom_segment(aes(y=position,yend=position, 
                              x=Start.date,xend=End.date,color=variant),
               size=3,alpha=1)+
  geom_text(aes(x=End.date+2,label=Description,y=position), fontface="bold", size=3,hjust=0)+ 
  geom_text(aes(x=Start.date-1,label=variant2,y=position), fontface="bold", size=4.5,hjust=1)+ 
  pubThemeDate+
  theme(axis.text.x=element_text(angle=0,hjust=0.5,size=9),
        axis.text.y=element_blank(),
        legend.position="none",
        axis.ticks.y=element_blank())+
  scale_x_date(breaks=xdates, date_labels = "%b %Y",
                            limits=c(as.Date(start.date)+11, as.Date(data.date)),
                            expand=expansion(add=c(14,260)))+  
  coord_cartesian(ylim=c(0.1,nrow(can.var2)))+
  scale_color_manual(values=respcolz)+
  scale_y_continuous(expand = expansion(add=c(0, 0)))+
  VOCColScale+
  #add annotation for restr1 and 2 for gamma
  annotate(geom="text",x=can.var2$Start.date[3],y=can.var2$position[3],
           label="restr.1",size=2.5,hjust=-.25,vjust=-1)+
  annotate(geom="text",x=can.var2$Start.date[4],y=can.var2$position[4],
           label="restr.2",size=2.5,hjust=.15,vjust=-1)

p2
ggsave(paste0(f.out,"Canada-variant-intervention-timeline.png"),height=1.7,width=8,units="in")

```

## Cartoon phylo
```{r}
variant.df.in<-paste0("colors_out/variant.df.can.csv")

## read in the variant df to use for inputs
var.df<-read.csv(variant.df.in)
var.df<-var.df[with(var.df, order(var.df$first.glob.date, decreasing = F)),]
excl.var<-c("Theta","Lambda","GH/490R","Other")
var.df<-var.df %>% dplyr::filter(!var.WHO %in% excl.var)
# var.df$var.WHO
var.df.phylo.order<-c("Beta","Epsilon","Eta","Iota","Mu","Kappa","Delta","Alpha","Gamma","Zeta","Omicron")
var.df$var.WHO<-factor(var.df$var.WHO, var.df.phylo.order)
var.df<-var.df[match(var.df.phylo.order,var.df$var.WHO),]
var.df$order<-1:nrow(var.df)
var.df$first.glob.date<-as.Date(var.df$first.glob.date)
var.df$first.can.date<-as.Date(var.df$first.can.date)

VOCColScale2<-scale_colour_manual(name = "Variant",values = varPalette.ch[which(names(varPalette.ch) %in% var.df.phylo.order)],na.value=NULL)
VOCFillScale2<-scale_fill_manual(name = "Variant",values = varPalette.ch[which(names(varPalette.ch) %in% var.df.phylo.order)],na.value=NULL)



# var.df$var.WHO<-factor(var.df$var.WHO,levels=var.ord)
# var.df
## RELAYY NEED TO CHECK THESE DATES

#delta missing first Can Date

var.df %>%
  ggplot()+
  geom_point(aes(x=first.can.date,y=order,color=var.WHO),size=6)+
  geom_text(aes(label=var.WHO,x=(first.can.date+10),y=order,color=var.WHO),hjust=-0.1)+
  VOCColScale2+
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none")+
    pubThemeDate+
  scaleDateFlex
ggsave(paste0(f.out,"firstVariantDetectionCanada.png"),height=5,width=5,units="in")

```


## make a composite plot
```{r}
plot_grid(p1,p2,align="v",nrow=2,rel_heights = c(0.4,0.1), labels=c("A","B"))
ggsave(paste0(f.out,"composite_InterventionsAndCases.png"),height=7,width=8.5,units="in")
```

### Want to add the clean Canadian sequences available over time with inferred dates
Also make a supplemental figure with the subsampling strategies and by lineage
DO THIS in the subsamp folder, easier to pick upthis whole script and plop in there to make a composite figure
