---
title: "merge partitions"
author: "Angela"
date: '2022-08-31'
output: html_document
---

# Objectives
* merge the partitioned metadata across clades
* little cleanup incl adding year-month, completing incomplete dates (sample day at random for canada meta)
* separate global and Canadian metadata
* make color schemes for VOCs and geos
* export metadata

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(stringr)
library(Biostrings)
library(lubridate)
library(gtools)
library(RColorBrewer)
library(stringi)
library(devtools)
library(grid)
library(gridExtra) 
```

## Ins and Outs
```{r}
data.date<-"2022-03-22" # the day of most recent sample

meta.in.folder<-"../00_cleanData/GISAID/metadata/03_cleaner"
meta.in.list<-list.files(meta.in.folder,full.names = T, pattern="cleaner.csv") %>% mixedsort()

#previously inferred dates for early Canadian sequences: take these as truth
meta.inferred.dates<-"../../20210301_phylogeo/03_ancestral/ft_ace_out/20210301_can100_1.meta.InferredDatesLSD.csv"

#metadata associated with reference sequence, wuhan-hu-1
wuhan.meta.in<-"../00_cleanData/GISAID/wuhan-hu-1/gisaid_hcov-19_2022_06_07_20_1.tsv"

#number of partitions
n.P<-length(meta.in.list)

#outputs, clean and separated can/global metadata
f.out<-"CleanMeta_PreSubsamp"
if(!dir.exists(f.out)){dir.create(f.out)}

meta.glob.out<-paste(f.out,"/global_meta.csv",sep="")
meta.can.out<-paste(f.out,"/canada_meta.csv",sep="")

#outputs, colors and lookup tables
f.out2<-"colors_out"
if(!dir.exists(f.out2)){dir.create(f.out2)} 

variant.colors.out<-paste(f.out2,"/variant_colors.tsv",sep="")
var.df.out<-paste(f.out2,"/variant.df.csv",sep="")
lookup.geo.out<-paste(f.out,"/lookup.geo.csv",sep="")
global.colors.out<-paste(f.out,"/global.colors.tsv",sep="")

#function to check colors
colortest<-function(vector){
  n<-length(vector)
  return(plot(1:n,1:n,col=vector,pch=15))
}

```

## Read in and link the metadata (in partitions)
```{r}
#list with element for each partition of gisaid clade
meta.l<-replicate(n.P,vector())

#clades
clades<-c()
for (k in 1:length(meta.in.list)){
  c1<-str_replace(last(unlist(str_split(meta.in.list[[k]],"/"))),"_cleaner.csv","")
  clades<-c(clades,c1)
}
names(meta.l)<-clades

for (k in 1:n.P){
  #META
  meta.l[[k]]<-read.csv(meta.in.list[k])

  #change a column name
  colnames(meta.l[[k]])[which(colnames(meta.l[[k]])=="Pango.lineage")]<-"Lineage"
  
  #dates 
  colnames(meta.l[[k]])[which(colnames(meta.l[[k]])=="Collection.date")]<-"date"
  
  #need to make region, country,division
  meta.l[[k]]<-tidyr::separate(meta.l[[k]],Location,sep="/",into=c("region","country","division"))
  
  #note that some have more or less than 3 locations, but should discard extra off the tail, or leave division blank
  #remove white space at beginning and end of location names
  meta.l[[k]]$region<-str_trim(meta.l[[k]]$region)
  meta.l[[k]]$country<-str_trim(meta.l[[k]]$country)
  #couple manual changes
  meta.l[[k]]$country<-str_replace_all(meta.l[[k]]$country,"USA\\? Ohio","USA")
  meta.l[[k]]$country<-str_replace_all(meta.l[[k]]$country,"Cote d\\'Ivoire","Cote dIvoire")
  meta.l[[k]]$country<-str_replace_all(meta.l[[k]]$country,"Usa","USA")
  meta.l[[k]]$division<-str_trim(meta.l[[k]]$division)

  #check no duplicate names
  length(meta.l[[k]]$new.names)==length(unique(meta.l[[k]]$new.names))

  #Remove any with only year (even if from Canada...)
  rem<-c()
  for (j in 1:nrow(meta.l[[k]])){
    if(length(unlist(strsplit(meta.l[[k]]$date[j],"-")))==1) {rem<-c(rem,j)}
  }
  if(length(rem)>1){
    meta.l[[k]]<-meta.l[[k]][-rem,]
  }

  #Remove any with date before 2019-12-01
  too.old<-which(meta.l[[k]]$date<as.Date("2019-12-01"))
  if(length(too.old)>1){
    meta.l[[k]]<-meta.l[[k]][-too.old,]
  }

  #also these buggers, for ex> 2020-01-XX
  if(any(str_detect(meta.l[[k]]$date,"-XX"))){
    meta.l[[k]]$date<-str_replace_all(meta.l[[k]]$date,"-XX","")
  }

  #make year-month a new column
  meta.l[[k]]$month<-as.character(NA)
  for (i in 1:nrow(meta.l[[k]])){
    if(length(unlist(strsplit(meta.l[[k]]$date[i],"-")))==2) {
      meta.l[[k]]$month[i]<-format(as.Date(paste(meta.l[[k]]$date[i],"-01",sep="")), "%Y-%m");next
      }
    if(length(unlist(strsplit(meta.l[[k]]$date[i],"-")))==3) {meta.l[[k]]$month[i]<-format(as.Date(meta.l[[k]]$date[i]), "%Y-%m")}
  }
}

## bind all the partitions together
meta<-bind_rows(meta.l)

## Add wuhan-hu-1 to metadata (not sure where it was excluded but not here)
wh1.1<-read.table(wuhan.meta.in,sep = "\t", header = T)
wh1.1$date<-wh1.1$Collection.date
wh1.1$country<-"China"
wh1.1$region<-"Asia"
wh1.1$division<-"Hubei"
wh1.1<-unite(wh1.1, "new.names",Accession.ID, date, sep="/",remove=F )
wh1.meta<-meta[1,]
wh1.meta[1,]<-NA
for (i in 1:ncol(wh1.meta)){
  match<-which(colnames(wh1.1) %in% colnames(wh1.meta)[i])
  if(length(match==1)){ wh1.meta[1,i]<-wh1.1[1,match]}
}
wh1.meta$month<-"2019-12"
wh1.meta$probability<-1

#BIND wuhan to meta
meta<-bind_rows(meta,wh1.meta)
```

## Separate Canadian metadata
```{r}
#subset the metadata
meta.can<-meta[which(meta$country=="Canada"),]
l.can<-nrow(meta.can) 

#check locations are right
# table(meta.can$division)
# meta.can$division[which(meta.can$division=="Toronto")]<-"Ontario"

# meta.can[which(is.na(meta.can$division)),] #UN-CRCHUM is in Montreal
if(any(is.na(meta.can$division))){
  meta.can$division[which(is.na(meta.can$division))]<-"Quebec"
}

#remove canada from meta for global subsample
meta<-meta[meta$country!="Canada",]

# Change to proper name
if(any(meta.can$division=="Newfoundland")){
  meta.can$division[which(meta.can$division=="Newfoundland")]<-"Newfoundland and Labrador"
}

```

## Deal with incomplete dates in Canada
```{r}
#Read in previously inferred dates from the 20210211 builds (some of which were inferred in 0115...)
meta.inf<-read.csv(meta.inferred.dates)

#Identify incomplete dates in Canada with new column 'incomp'
#If available in previously inferred dates, replace
#Else, sample at random
up.date<-c()
meta.can$incomp<-0
for (i in 1:nrow(meta.can)){
  dt<-unlist(strsplit(meta.can$date[i],split="-"))
  if(length(dt)==2){
  #it is incomplete
  meta.can$incomp[i]<-1
  #look for a match in the prev'ly inferred
  matchy<-which(meta.inf$GISAID_ID %in% meta.can$Accession.ID[i])
  if(length(matchy)>0){
    meta.can$date[i]<-meta.inf$date.lsd[matchy]
    up.date<-c(up.date,i) #add to running vector of replacements
    next #skip to next i
  }
  #Else, find days in month
  dayz<-lubridate::days_in_month(paste(dt[1],dt[2],"01",sep="-")) %>% as.integer
  #Then, sample a day and plop it in the date
  day.samp<-formatC(sample(dayz,size=1), width = 2, format = "d", flag = "0") #make sure day has trailing zero
  meta.can$date[i]<-paste(dt[1],dt[2],day.samp,sep="-") 
  }
}

#how many prev inferred?
length(up.date)

#look at incompletes
# table(meta.can$incomp)
# 31526/nrow(meta.can) #11% cool...
# table(meta.can$incomp, meta.can$division) #ontario contributed 31525 incomplete dates
```

## Add columns for the variant into the meta
```{r}
#group lineages within variants (grouped by gisaid)
# sort(table(meta.can$Lineage))
var.df<-data.frame("Variant"=names(table(meta.can$Variant)))

#split the string by spaces
var.df<-var.df %>% separate(Variant, sep=" ", as.character(1:10),remove=F)

#quick change for last field
var.df[nrow(var.df),5:9]<-var.df[nrow(var.df),4:8]
var.df[nrow(var.df),4]<-var.df[nrow(var.df),3]
var.df<-var.df %>% unite("origin","5":"10",sep=" ",na.rm = T)
colnames(var.df)<-c("Variant","var.type","var.WHO","var.clade","var.lin","var.origin")
var.df[1,c(2,3)]<-"Other"
var.df$var.lin<-str_replace_all(var.df$var.lin,"\\(|\\)","")
#add first global date from covlineages.org
var.df$first.glob.date<-c("2019-12-24", "2020-09-01", "2020-09-01", "2021-03-01",
                         "2020-09-22","2021-09-02", "2020-04-11","2020-03-25",
                         "2020-01-28", "2020-03-03","2020-07-21","2020-12-15",
                         "2021-01-08","2020-04-13","2021-01-01")

### ORDER by greek
# c("Alpha", "Beta", "Gamma", "Delta", "Epsilon", "Zeta", "Eta", "Iota", "Kappa", "Mu","Omicron")
# See the _mod version of var.df for updated (as of 2023-05-11) first global sample dates

## ORDER by Canada
## ORDER by global

#order by their first global date
var.df<-var.df[with(var.df, order(as.Date(var.df$first.glob.date), decreasing = F)),]
write.csv(var.df,var.df.out,row.names = F)
#give each one a color
vars.ch<-var.df$var.WHO
varcolz<-colorRampPalette(c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#ffff99','#b15928'))(nrow(var.df)) #https://colorbrewer2.org/#type=qualitative&scheme=Paired&n=12
names(vars.ch)<-varcolz

#check
vars.ch.df<-as.data.frame(vars.ch)
ggplot(vars.ch.df)+
  geom_text(aes(x=1:15, y=1:15,label=vars.ch, color=vars.ch))+
  scale_color_manual(values=row.names(vars.ch.df))

#write the color scheme for future use
write.table(vars.ch.df,variant.colors.out,sep="\t",row.names=T)
```

##Join variant df onto meta
```{r}
#find rows with variant==na
meta$Variant<-na.replace(meta$Variant,)
meta.can$Variant<-na.replace(meta.can$Variant, "")

#add the var.df to to meta and metacan
meta<-left_join(meta, var.df, by="Variant")
meta.can<-left_join(meta.can, var.df, by="Variant")

#check if all have var.WHO 
any(is.na(meta$var.WHO))
any(is.na(meta.can$var.WHO))
any(meta$var.WHO=="")
any(meta.can$var.WHO=="")

#remove unnecessary columns
meta<-meta[,-which(colnames(meta) %in% c("X","X.1","probability"))]
meta.can<-meta.can[,-which(colnames(meta.can) %in% c("X","X.1","probability"))]
print(nrow(meta))
print(nrow(meta.can))

#weirdo:
bady<-which(meta.can$date=="2021")
if(length(bady)>0){
  meta.can<-meta.can[-bady,]
}
```

# Write the merged and cleaned data
```{r}
#fully compiled and cleaned meta + metacan
write.csv(meta,meta.glob.out,row.names = F)
write.csv(meta.can,meta.can.out,row.names=F)
```

# Make colors and lookups for global regions, Canadian provinces
```{r}
## Make a lookup table of locations, grouped if sparser sequences
#build lookup table
allcountries<-unique(meta.glob$country) 
length(allcountries)#210
lookup.geo<-data.frame(og.loc=allcountries,
                       new.loc=NA)

#identify all 'big.origins' countries that contributed more than the 95th percentile of seqs
# summary(as.vector(table(meta.glob$country)))
seq.count<-table(meta.glob$country)
# summary(as.vector(seq.count))
#median 721.0 ; mean 39577.3 
perc90<-quantile(as.vector(seq.count),probs = 0.9)
perc95<-quantile(as.vector(seq.count),probs = 0.95)
#use 95th percentile
big.origins<-c(names(seq.count[which(seq.count>perc95)]))

#add on south africa due to importance in beta and omicron
big.origins<-c(big.origins,"South Africa")
# length(big.origins) #12

#go through all countries, if a big.origin, keep the name. otherwise use region/continent as lookup geo
for (i in 1:nrow(lookup.geo)){
  ifelse (lookup.geo$og.loc[i] %in% big.origins,
          lookup.geo$new.loc[i]<-lookup.geo$og.loc[i],
          lookup.geo$new.loc[i]<-meta.glob$region[which(meta.glob$country==lookup.geo$og.loc[i])[1]]
  )
}
# head(lookup.geo)

#add provinces to lookup.geo and keep original, no maritimes merging
allprovs<-unique(meta.can$division) 
lookup.prov<-data.frame(og.loc=allprovs,
                        new.loc=allprovs)
#merge the dfs
lookup.geo.final<-lookup.geo %>% bind_rows(lookup.prov)

#write the lookuptable for future use
write.csv(lookup.geo.final,lookup.geo.out,row.names=F)

##### LOCATION COLOR SCHEMES ######

#first pick out divergent color scheme for provinces
provcols<-c("#486090","plum4","brown4","brown3","darkorange1","darkorange3",
            "darkgoldenrod1","darkgoldenrod2","darkgoldenrod3","darkgoldenrod4", 
             "bisque1","bisque2","bisque4")
provs<-c("Quebec","British Columbia","Ontario","Alberta","Manitoba","Saskatchewan",
         "New Brunswick","Newfoundland and Labrador","Nova Scotia","Prince Edward Island",
         "Yukon","Northwest Territories","Nunavut")
names(provcols)<-as.character(provs)
colortest(provcols)

#assign continents and their big.origins countries grouped color schemes
#https://colorbrewer2.org/#type=sequential&scheme=YlGn&n=9
asia.cols<-c("#005a32","#41ab5d","#addd8e") #greens, dark to light
names(asia.cols)<-c("Asia","India","Japan")
colortest(asia.cols)

#https://colorbrewer2.org/#type=sequential&scheme=YlGnBu&n=9
eur.cols<-rev(c('#edf8b1','#c7e9b4','#7fcdbb','#41b6c4','#1d91c0','#225ea8','#253494','#081d58'))
names(eur.cols)<-c("Europe","France","United Kingdom","Germany","Spain","Sweden","Switzerland","Denmark")
colortest(eur.cols)

#https://colorbrewer2.org/#type=sequential&scheme=Purples&n=7
amer.cols<-rev(c('#9e9ac8','#807dba','#6a51a3','#4a1486'))
names(amer.cols)<-c("North America","USA","South America","Brazil")
colortest(amer.cols)

#single color to represent canada DARK RED
can.cols<-"#8b0000"
names(can.cols)<-"Canada"

#https://colorbrewer2.org/#type=sequential&scheme=RdPu&n=7
afr.cols<-rev(c("#f768a1","#ae017e")) #pink
names(afr.cols)<-c("Africa","South Africa")
colortest(afr.cols)

oce.cols<-c("#fa9fb5") #light pink
names(oce.cols)<-"Oceania"
colortest(oce.cols)

#Add all together
globalPalette<-c(provcols,can.cols,eur.cols,amer.cols,asia.cols,afr.cols,oce.cols)
# nm<-names(globalPalette)
# globalPalette<-as.data.frame(globalPalette)
# rownames(globalPalette)<-nm

#example plot
colortest(globalPalette)

#export a tsv of name and hex color for others to use
write.table(as.data.frame(globalPalette),global.colors.out,sep="\t",
            row.names = T)

```

