---
title: "corr coef"
author: "Angela"
date: '2022-10-19'
output: html_document
---

## Objectives
* compare the correlation coefficients between cases and sequences for linear and log, for Canada and global, for the following:
  * pre-subsampling
  * post-subsampling, temporal distrib
  * post-subsampling, uniform distrib
  
```{r}
library(ggplot2)
library(tidyverse)
```

```{r}
pubTheme<-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background=element_rect("grey95"), axis.line = element_line(colour = "black"),
            legend.key.size = unit(0.5,"line"),
            text=element_text(size=10,face="bold"),
            legend.text=element_text(size=6))

f.out<-"compare_subsamps_2023/"
if(!dir.exists(f.out)) {dir.create(f.out)}
```
  
  
```{r}
# pre-subsampling
pre.can.lin<-read.csv("Results_PreSubsamp_omi_2023/corr.case.vs.seq.Canada.pre-samp.csv")
pre.can.log<-read.csv("Results_PreSubsamp_omi_2023/corr.case.vs.seq.Canada.pre-samp.log10.csv")
pre.glob.lin<-read.csv("Results_PreSubsamp_omi_2023/corr.case.vs.seq.global.pre-samp.csv")
pre.glob.log<-read.csv("Results_PreSubsamp_omi_2023/corr.case.vs.seq.global.pre-samp.log10.csv")

#post-subsampling, temporal distrib 1 (sampling probabilities based on raw instead of rolling avg estimated counts)
temp1.can.lin<-read.csv("Results_PostSubsamp_202208_2023/corr.case.vs.seq.Canada.post-samp.csv")
temp1.can.log<-read.csv("Results_PostSubsamp_202208_2023/corr.case.vs.seq.Canada.post-samp.log10.csv")
temp1.glob.lin<-read.csv("Results_PostSubsamp_202208_2023/corr.case.vs.seq.global.post-samp.csv")
temp1.glob.log<-read.csv("Results_PostSubsamp_202208_2023/corr.case.vs.seq.global.post-samp.log10.csv")

# #post-subsampling, temporal distrib 2 - first bootstrap (sampling probabilities based on rolling avg estimated counts)
# temp2.1.can.lin<-read.csv("Results_PostSubsamp_202210/corr.case.vs.seq.Canada.post-samp.csv")
# temp2.1.can.log<-read.csv("Results_PostSubsamp_202210/corr.case.vs.seq.Canada.post-samp.log10.csv")
# temp2.1.glob.lin<-read.csv("Results_PostSubsamp_202210/corr.case.vs.seq.global.post-samp.csv")
# temp2.1.glob.log<-read.csv("Results_PostSubsamp_202210/corr.case.vs.seq.global.post-samp.log10.csv")

#post-subsampling, temporal distrib 2 - second bootstrap (sampling probabilities based on rolling avg estimated counts)
# temp2.2.can.lin<-read.csv("Results_PostSubsamp_202210_2/corr.case.vs.seq.Canada.post-samp.csv")
# temp2.2.can.log<-read.csv("Results_PostSubsamp_202210_2/corr.case.vs.seq.Canada.post-samp.log10.csv")
# temp2.2.glob.lin<-read.csv("Results_PostSubsamp_202210_2/corr.case.vs.seq.global.post-samp.csv")
# temp2.2.glob.log<-read.csv("Results_PostSubsamp_202210_2/corr.case.vs.seq.global.post-samp.log10.csv")

# post-subsampling, uniform distrib
unif.can.lin<-read.csv("Results_PostSubsamp_unif_2023/corr.case.vs.seq.Canada.post-samp.csv")
unif.can.log<-read.csv("Results_PostSubsamp_unif_2023/corr.case.vs.seq.Canada.post-samp.log10.csv")
unif.glob.lin<-read.csv("Results_PostSubsamp_unif_2023/corr.case.vs.seq.global.post-samp.csv")
unif.glob.log<-read.csv("Results_PostSubsamp_unif_2023/corr.case.vs.seq.global.post-samp.log10.csv")

# post-subsampling, uniform distrib
# unif.can.lin<-read.csv("Results_PostSubsamp_unif/corr.case.vs.seq.Canada.post-samp.csv")
# unif.can.log<-read.csv("Results_PostSubsamp_unif/corr.case.vs.seq.Canada.post-samp.log10.csv")
# unif.glob.lin<-read.csv("Results_PostSubsamp_unif/corr.case.vs.seq.global.post-samp.csv")
# unif.glob.log<-read.csv("Results_PostSubsamp_unif/corr.case.vs.seq.global.post-samp.log10.csv")
```

## add descriptor columns to each
```{r}
add.strat<-function(df,strat,type,geo){
  #add descriptor cols
  df$strategy<-strat
  df$type<-type
  df$geo<-geo
  return(df)
}

pre.can.lin<-add.strat(pre.can.lin,"pre-samp","linear","canada")
pre.can.log<-add.strat(pre.can.log,"pre-samp","log","canada")
pre.glob.lin<-add.strat(pre.glob.lin,"pre-samp","linear","global")
pre.glob.log<-add.strat(pre.glob.log,"pre-samp","log","global")

#post-subsampling, temporal distribs
temp1.can.lin<-add.strat(temp1.can.lin,"temp1","linear","canada")
temp1.can.log<-add.strat(temp1.can.log,"temp1","log","canada")
temp1.glob.lin<-add.strat(temp1.glob.lin,"temp1","linear","global")
temp1.glob.log<-add.strat(temp1.glob.log,"temp1","log","global")

# temp2.1.can.lin<-add.strat(temp2.1.can.lin,"temp2","linear","canada")
# temp2.1.can.log<-add.strat(temp2.1.can.log,"temp2","log","canada")
# temp2.1.glob.lin<-add.strat(temp2.1.glob.lin,"temp2","linear","global")
# temp2.1.glob.log<-add.strat(temp2.1.glob.log,"temp2","log","global")
# 
# temp2.2.can.lin<-add.strat(temp2.2.can.lin,"temp2.2","linear","canada")
# temp2.2.can.log<-add.strat(temp2.2.can.log,"temp2.2","log","canada")
# temp2.2.glob.lin<-add.strat(temp2.2.glob.lin,"temp2.2","linear","global")
# temp2.2.glob.log<-add.strat(temp2.2.glob.log,"temp2.2","log","global")

# post-subsampling, uniform distrib
unif.can.lin<-add.strat(unif.can.lin,"uniform","linear","canada")
unif.can.log<-add.strat(unif.can.log,"uniform","log","canada")
unif.glob.lin<-add.strat(unif.glob.lin,"uniform","linear","global")
unif.glob.log<-add.strat(unif.glob.lin,"uniform","log","global")
```

#join them all together
```{r}
all.lin<-bind_rows(pre.can.lin, pre.glob.lin, temp1.can.lin, temp1.glob.log,
                   #temp2.1.can.lin, temp2.1.glob.lin, temp2.2.can.lin, temp2.2.glob.lin,
                   unif.can.lin, unif.glob.lin)
all.log<-bind_rows(pre.can.log, pre.glob.log, temp1.can.log, temp1.glob.log,
                   #temp2.1.can.log, temp2.1.glob.log, temp2.2.can.log, temp2.2.glob.log,
                   unif.can.log, unif.glob.log)

#order the lineages 
var.ord<-c("Other","Alpha","Zeta","Gamma", "Epsilon","Eta","Beta","Delta",
           "Kappa","Iota","Mu","BA.1","BA.1.1","BA.2")
all.lin$var.WHO<-factor(all.lin$var.WHO,levels=var.ord)
all.log$var.WHO<-factor(all.log$var.WHO,levels=var.ord)

#order the strats
strat.ord<-c("pre-samp","uniform","temp1")#,"temp2.1","temp2.2")
all.lin$strategy<-factor(all.lin$strategy,levels=strat.ord)
all.log$strategy<-factor(all.log$strategy,levels=strat.ord)
```

#Plot the pearson and spearman correlation coeffs for linear case to seq
```{r}
#compare the two boots
all.lin %>%
  # filter(strategy %in% c("temp2.1","temp2.2")) %>%
  ggplot()+
  geom_point(aes(x=geo, y=pearson, color=strategy),alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="free_y")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+
  labs(x=NULL)+
  pubTheme
ggsave(paste0(f.out,"temporal2_2boots_linear_pearson.png"),width=6,height=3,units="in")
#almost identical, except for other

all.lin %>%
  filter(strategy %in% c("temp2.1","temp2.2")) %>%
  ggplot()+
  geom_point(aes(x=geo, y=spearman, color=strategy),alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="free_y")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+
  labs(x=NULL)+
  pubTheme
ggsave(paste0(f.out,"temporal2_2boots_linear_spearman.png"),width=6,height=3,units="in")

#interesting, they vary more for the spearman rank correlation... would expect this to be more consistent? maybe to do with excluding zeroes?
#should go back and repeat this export ONLY for all boots of the desired strategy

#compare strats, remove one boot for now
all.lin.mod<-all.lin
library(stringi)
all.lin.mod$strategy<- stri_replace_all_regex(all.lin.mod$strategy,
                                  pattern=c('pre-samp', 'temp1', 'temp2.1',"temp2.2","uniform"),
                                  replacement=c('Pre-subsample', 'Temporal-sampling', 'Temporal2.1',"Temporal2.2","Uniform-sampling"),
                                  vectorize=FALSE)

all.lin.mod %>%
  filter(!strategy %in% c('Temporal2.1',"Temporal2.2")) %>%
  ggplot()+
  pubTheme+
  geom_jitter(aes(x=geo, y=pearson, color=strategy),stat="identity",width=0.05,alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="free_y")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1),
        legend.position = "top")+
  labs(x=NULL, y="Pearson r: monthly cases vs. sequences")+
  guides(color=guide_legend(title="Sample", title.position = "left"))+
  scale_y_continuous(expand=expansion(0.05,0.05))
ggsave(paste0(f.out,"fullcomp_linear_pearson.png"),width=6,height=3,units="in")

#compare strats, remove one boot for now
all.lin.mod %>% 
  filter(!strategy %in% c('Temporal2.1',"Temporal2.2")) %>%
  ggplot()+
  geom_jitter(aes(x=geo, y=spearman, color=strategy),stat="identity",width=0.05,alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="free_y")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+
  labs(x=NULL, y="Spearman rank corr. coeff. cases vs sequences")+
  pubTheme+
  scale_y_continuous(expand=expansion(0.05,0.05))
ggsave(paste0(f.out,"fullcomp_linear_spearman.png"),width=6,height=3.5,units="in")

#compare R2
all.lin.mod %>% 
  filter(!strategy %in% c('Temporal2.1',"Temporal2.2")) %>%
  ggplot()+
  geom_jitter(aes(x=geo, y=adj.r2, color=strategy),stat="identity",width=0.05,alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="free_y")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+
  labs(x=NULL, y="Adjusted R^2, linear fit cases v seqs")+
  pubTheme+
  scale_y_continuous(expand=expansion(0.05,0.05))
ggsave(paste0(f.out,"fullcomp_linear_R2.png"),width=6,height=3.5,units="in")

```


#Plot the pearson and spearman correlation coeffs for linear case to seq
```{r}
#compare the two boots
all.log %>%
  filter(strategy %in% c("temp2.1","temp2.2")) %>%
  ggplot()+
  geom_point(aes(x=geo, y=pearson, color=strategy),alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="free_y")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1),
        legend.position = "top")+
  labs(x=NULL, y="Pearson r: monthly cases vs. sequences")+
  guides(color=guide_legend(title="Sample", title.position = "left"))+
  scale_y_continuous(expand=expansion(0.05,0.05))
ggsave(paste0(f.out,"temporal2_2boots_log_pearson.png"),width=6,height=3,units="in")

all.log %>%
  filter(strategy %in% c("temp2.1","temp2.2")) %>%
  ggplot()+
  geom_point(aes(x=geo, y=spearman, color=strategy),alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="fixed")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1),
        legend.position = "top")+
  labs(x=NULL, y="Spearman r: monthly cases vs. sequences")+
  guides(color=guide_legend(title="Sample", title.position = "left"))+
  scale_y_continuous(expand=expansion(0.05,0.05))
ggsave(paste0(f.out,"temporal2_2boots_log_spearman.png"),width=6,height=3,units="in")

#should go back and repeat this export ONLY for all boots of the desired strategy

#compare strats, remove one boot for now
all.log.mod<-all.log
all.log.mod$strategy<- stri_replace_all_regex(all.log.mod$strategy,
                                  pattern=c('pre-samp', 'temp1', 'temp2.1',"temp2.2","unif"),
                                  replacement=c('Pre-subsample', 'Temporal-sampling', 'Temporal2.1',"Temporal2.2","Uniform-sampling"),
                                  vectorize=FALSE)


all.log.mod %>%
  filter(!strategy %in% c('Temporal2.1',"Temporal2.2")) %>%
  ggplot()+
  geom_jitter(aes(x=geo, y=pearson, color=strategy),stat="identity",width=0.05,alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="free_y")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1),
                legend.position = "top")+
  pubTheme+
  labs(x=NULL, y="Pearson r: log monthly cases vs. sequences")+
  guides(color=guide_legend(title="Sample", title.position = "left"))+
  scale_y_continuous(expand=expansion(0.05,0.05))
ggsave(paste0(f.out,"fullcomp_log_pearson.png"),width=6,height=3,units="in")

#compare strats, remove one boot for now
all.log.mod %>%
  filter(!strategy %in% c('Temporal2.1',"Temporal2.2")) %>%
  ggplot()+
  geom_jitter(aes(x=geo, y=spearman, color=strategy),stat="identity",width=0.05,alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="free_y")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1),
                legend.position = "top")+
  pubTheme+
  labs(x=NULL, y="Spearman r: log monthly cases vs. sequences")+
  guides(color=guide_legend(title="Sample", title.position = "left"))+
  scale_y_continuous(expand=expansion(0.05,0.05))
ggsave(paste0(f.out,"fullcomp_log_spearman.png"),width=6,height=3,units="in")

all.log.mod %>%
  filter(!strategy %in% c('Temporal2.1',"Temporal2.2")) %>%
  ggplot()+
  geom_jitter(aes(x=geo, y=adj.r2, color=strategy),stat="identity",width=0.05,alpha=0.9)+
  facet_wrap(~var.WHO,nrow=3,scales="free_y")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1),
                legend.position = "top")+
  pubTheme+
  labs(x=NULL, y="Adjusted R^2: linear fit log cases vs. seqs")+
  guides(color=guide_legend(title="Sample", title.position = "left"))+
  scale_y_continuous(expand=expansion(0.05,0.05))
ggsave(paste0(f.out,"fullcomp_log_r2.png"),width=6,height=3,units="in")

```


# Calculate and plot the difference in coefficients (pre to post)
```{r}
## what about looking at diffs 
```

  