---
title: "Untitled"
author: "Angela"
date: '2023-06-26'
output: html_document
---


# Setup

## libraries
```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(stringr)
library(coronavirus)
library(ggplot2)
library(RColorBrewer)
library(zoo)
library(ggstance)
library(cowplot)
library(lubridate)
library(MASS)
library(ggrepel)
library(gtools)
library(data.table)
library(ggplotify)
library(gridExtra)
library(grid)
library(lattice)
library(reshape2)
library(gdata) 
```

## inputs and outputs
```{r}
# the day of most recent sample
data.date<-"2022-03-22" 
start.date<-"2020-11-01" #about when VOCs first started coming to Canada

# set up months for dataframes below for 2021 and 2022
last.mon<-format(as.Date(data.date), "%m")
last.mon.l<-24+as.numeric(last.mon) #starts at 0, so actually leads to 13+..
mons <- format(ymd(as.Date("2019-12-01")) %m+% months(0:last.mon.l),"%Y-%m")
l.mons<-length(mons)

#pre-subsampled clean data for global and canada
meta.in.glob<-"CleanMeta_PreSubsamp/global_meta.csv"
meta.in.can<-"CleanMeta_PreSubsamp/canada_meta.csv" 

## Geo lookup table and colors from previous build
lookup.geo.in<-"colors_out/lookup.geo.csv"
geo.colors.in<-"colors_out/global.colors.tsv"

#continents
continents.in<-"../../20220322_phylogeo/01_subSample/CleanMeta_PreSubsamp/countries.csv"

#variant colors
variant.colors.in<-"CleanMeta_PreSubsamp/variant_colors.tsv"
variant.df.in<-"CleanMeta_PreSubsamp/variant.df.csv" #couple manual updates 2023-05-11 based on outbreak.info


f.in<-"Results_PreSubsamp_omi_2023"

#kept omicron the same even though it is 11 b/c plausilbe that can seqs before detected 

#PHAC cases over time by prov
can.cases.in<-"Canada_cases/20220516_covid19_casesbyprov.csv"

#folder out
f.out<-"Figure1_plot_20230731"
if(!dir.exists(f.out)){dir.create(f.out)}

## Setup themes for plots
pubTheme<-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            panel.background=element_rect("grey95"),
            axis.line = element_line(colour = "black"),
            legend.key.size = unit(0.5,"line"),
            text=element_text(size=10,face="bold"),
            legend.text=element_text(size=8))

pubThemeDate<-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    panel.background = element_rect("grey95"), 
                    axis.line = element_line(colour = "black"),
                    text=element_text(size=10,face="bold"),
                    legend.text=element_text(size=8),
                    axis.text.x=element_text(angle = 45,hjust=1,size=rel(1))) #this is the only difference b/w them

scaleDate<-scale_x_date(date_breaks = "1 month", date_minor_breaks = "2 weeks", date_labels = "%b %Y",limits=as.Date(c(start.date,"2022-03-22")),expand=c(0,0))

xdates<-ymd(c("2020-12-01","2021-03-01","2021-06-01","2021-09-01","2021-12-01","2022-03-01"))
scaleDateBusy<-scale_x_date(breaks=xdates, date_labels = "%b %Y",
                            limits=as.Date(c(start.date,"2022-03-22")),
                            expand=c(0,0))

scaleDateFlex<-scale_x_date(date_labels = "%b %Y",expand=c(0,0))

scaleMonthChar<-scale_x_discrete(breaks=c("2020-12","2021-03","2021-06","2021-09","2021-12","2022-03"),labels=c("Dec 2020","Mar 2021","Jun 2021","Sep 2021","Dec 2021","Mar 2022"))

```

## Make a color scheme for all countries with seqs
```{r}
#Use new color scheme with fewer groupings
lookup.geo<-read.csv(lookup.geo.in,header=T) #use later on

#import a tsv of name and hex color for LOCATIONS
globalPalette<-read.table(geo.colors.in,sep="\t")

## make color scheme
glob.colz<-row.names(globalPalette)
globalPalette.ch<-as.character(globalPalette$globalPalette)
names(globalPalette.ch)<-glob.colz

GlobColScale<-scale_colour_manual(name = "Location",values = globalPalette.ch,na.value="grey60")
GlobFillScale<-scale_fill_manual(name = "Location",values = globalPalette.ch,na.value="grey60")


CountryColScale<-scale_colour_manual(name = "Global region",
                                     values = globalPalette.ch[15:length( globalPalette.ch)],
                                     na.value="grey60")
CountryFillScale<-scale_fill_manual(name = "Global region",
                                    values = globalPalette.ch[15:length( globalPalette.ch)]
                                    ,na.value="grey60")
```

## VOC color scheme
Note this is read in from the 00_MergePartitionsSplitCanada script
```{r}
varPalette<-read.table(variant.colors.in,sep="\t")

## make color scheme
var.colz<-as.character(varPalette$vars.ch)
varPalette.ch<-row.names(varPalette)
names(varPalette.ch)<-var.colz

# var.colz<-row.names(varPalette)
# varPalette.ch<-as.character(varPalette$vars.ch)
# names(varPalette.ch)<-var.colz

VOCColScale<-scale_colour_manual(name = "Variant",values = varPalette.ch,na.value=NULL)
VOCFillScale<-scale_fill_manual(name = "Variant",values = varPalette.ch,na.value=NULL)

### FOR SEPEARATE OMI###
### add separate groups for each omicron group
omis<-c("BA.1","BA.1.1","BA.2", "BA.3")
# https://www.colorhexa.com/ec9a91
omiPalette.ch<-c("#e47266","#e8867b","#f0aea7","#f4c2bc") 
names(omiPalette.ch)<-omis
morevarPal<-c(varPalette.ch,omiPalette.ch)
morevarPal<-morevarPal[-which(names(morevarPal)=="Omicron")]

VarColScale<-scale_colour_manual(name = "Variant",values = morevarPal,na.value=NULL)
VarFillScale<-scale_fill_manual(name = "Variant",values = morevarPal,na.value=NULL)

## read in the variant df to later remove metacan seqs with dates that are too early
# var.df<-read.csv(variant.df.in)
# 
# #add to var.df
# var.omi<-var.df[which(var.df$var.WHO=="Omicron"),]
# var.omi<-rbind(var.omi,var.omi,var.omi,var.omi) #one for each sep lineage considered
# var.omi$var.WHO<-omis
# var.omi$var.lin<-omis
# var.omi$first.glob.date<-c("2021-09-02","2021-09-09","2021-09-21","2021-11-23")
# # var.omi$first.can.date<-c("2021-11-04","2021-11-04","2021-11-29")
# 
# #join var.omi onto var.df 
# var.df<-rbind(var.df,var.omi)
# 
# #remove 'omicron' full group
# var.df<-var.df[-which(var.df$var.WHO=="Omicron"),]
# 

var.df<-read.csv(paste0(f.in,"/variant.df.can.csv"))
```



## Read in variant cases
```{r}
meta.can.var.daily.full3<-read.csv(paste0(f.in,"/meta.can.var.daily.full3.csv"))

```

##Make a version of the variant cases in Canada with BA.* separate
```{r}

#order the variants   
var.df<-var.df[with(var.df, order(var.df$first.can.date, decreasing = F)),]
var.df<-var.df[c(1:14,16,15,17),] #make ba.1 come before ba.1.1
var.ord<-var.df$var.WHO
var.ord.red<-var.ord[-which(var.ord %in% c("Theta","Lambda","GH/490R","BA.3"))]
var.ord.red2<-var.ord.red
#merge BA.1 iwth BA.1.1:
# var.ord.red2<-var.ord.red[-which(var.ord.red %in% c("BA.1.1"))]
# var.ord.red2<-var.ord.red[-which(var.ord.red %in% c( "BA.1","BA.1.1" ,"BA.2"))]
# var.ord.red2<-var.ord.red2 %>% c("Omicron")

morevarPal2<-morevarPal
# morevarPal2<-morevarPal2[-which(names(morevarPal2) %in% c( "BA.1.1"))]
# names(morevarPal2)[which(names(morevarPal2)=="BA.1.1")]<-"Omicron"

VarFillScaleRed2<-scale_fill_manual(name = "Variant",values = morevarPal2[match(var.ord.red2,names(morevarPal2))],na.value="grey60")
VarColScaleRed2<-scale_color_manual(name = "Variant",values = morevarPal2[match(var.ord.red2,names(morevarPal2))],na.value="grey60")


meta.can.var.daily.full3.2<-meta.can.var.daily.full3
# meta.can.var.daily.full3.2$var.WHO<-str_replace_all(meta.can.var.daily.full3.2$var.WHO, c("BA.1.1"="BA.1"))
# meta.can.var.daily.full3.2$var.WHO<-str_replace_all(meta.can.var.daily.full3.2$var.WHO,"Omicron.1","Omicron")
meta.can.var.daily.full3.2$var.WHO<-factor(meta.can.var.daily.full3.2$var.WHO,
                                           levels=var.ord.red2)
meta.can.var.daily.full3.2$date<-as.Date(meta.can.var.daily.full3.2$date)

#Plot the avgVOC_cases by var.WHO for all of Canada
plot.daily.ave.voc.cases.f1<-meta.can.var.daily.full3.2%>%
  ggplot()+
  pubThemeDate+
  theme(axis.title.y=element_text(size=rel(1)),
        legend.position=c(0.3,0.6),
        legend.text=element_text(size=rel(0.9)),
        legend.title = element_text(size=rel(1.1)),
        plot.margin=margin(1,1,1,10,unit="pt"),
        axis.text.x=element_blank())+
  geom_bar(aes(x=date,y=avgVOC_cases,fill=var.WHO),stat="identity",width=1)+
  labs(x=NULL, y="Daily variant cases in Canada")+
  scale_y_continuous(expand=c(0,0))+
  VarFillScaleRed2+
  scaleDateBusy+
  guides(fill=guide_legend(title.position = "top",keywidth = 1,keyheight = 1, 
                           title="Variant",ncol=2))
  
plot.daily.ave.voc.cases.f1
ggsave(paste(f.out,"/AverageDailyVOCSpecificCases_Canada_FIG1A.png",sep=""), height=3,width=5,units="in")

#make a version with waves labelled
lb2 <- paste("2^nd","~wave")
lb3 <- paste("3^rd","~wave")
lb4 <- paste("4^th","~wave")
lb5 <- paste("5^th","~wave")

plot.daily.ave.voc.cases.fplus<-plot.daily.ave.voc.cases.f1+
  annotate(geom="text",label=lb2,parse=TRUE,fontface="italic", 
           x=as.Date("2021-01-05"), y=8400, hjust=1,size=2.5)+
  annotate(geom="text",label=lb3,parse=TRUE,fontface="italic", 
           x=as.Date("2021-04-09"), y=8400, hjust=1,size=2.5)+
  annotate(geom="text",label=lb4,parse=TRUE,fontface="italic", 
           x=as.Date("2021-09-15"), y=5050, hjust=1,size=2.5)+
  annotate(geom="text",label=lb5,parse=TRUE,fontface="italic", 
           x=as.Date("2021-12-28"), y=40000, hjust=1,size=2.5)
plot.daily.ave.voc.cases.fplus
ggsave(paste(f.out,"/AverageDailyVOCSpecificCases_Canada_FIG1A_waves.png",sep=""), height=3,width=5,units="in")

```

```{r}
interv<-read.csv("../04_canadaTimeline/2023-02-17_Results/can.variant.interv.plotty2.csv")
interv$Start.date<-as.Date(interv$Start.date)
interv$End.date<-as.Date(interv$End.date)
#Make more concise
interv$Description[interv$variant=="Delta"]<-"India flight ban"
interv$Description[interv$variant=="Alpha"]<-"UK flight ban"
interv$Description[interv$variant=="Beta"]<-"Enhanced screening South Africa"
interv$Description[interv$variant=="Gamma"]<-"Enhanced screening Brazil"
interv$Description[interv$variant=="Omicron"]<-"Southern Africa entry ban"

can.var2<-interv
#make y positions (same for two gamma interventions)
can.var2$position<-c(1,2,3,3,4,5)
#bigger gaps for larger scale
can.var2$position2<-can.var2$position*150
can.var2$hjusts<-c(0,0,0,0,0,0.85)
# can.var2$position<-1:nrow(can.var2)

#remove redundant text for two gamma interves
can.var2$Description[4]<-NA
can.var2$variant2<-can.var2$variant #duplicate
can.var2$variant2[4]<-NA
can.var2$var.WHO<-can.var2$variant

#replace the color name
can.var2$var.WHO<-str_replace_all(can.var2$var.WHO, "Omicron","BA.1")

# can.var2$variant[3]<-NA #need this text 
# head(can.var)
# p2<-can.var2 %>%
#   ggplot()+
#   labs(x=NULL,y=NULL)+
#   #description and lines to axis
#   geom_segment(aes(y=position,yend=position, 
#                               x=Start.date,xend=End.date,color=variant),
#                size=3,alpha=1)+
#   geom_text(aes(x=Start.date,label=Description,y=position+0.25), fontface="bold", size=3,hjust=0)+
#   # geom_text(aes(x=Start.date-1,label=variant2,y=position), fontface="bold", size=4.5,hjust=1)+ 
#   pubThemeDate+
#   theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1),
#         axis.text.y=element_blank(),
#         legend.position="none",
#         axis.ticks.y=element_blank(),
#         )+
#   # scaleDateBusy+
#   scale_x_date(breaks=xdates, date_labels = "%b %Y",
#                             limits=c(as.Date(start.date)+11, as.Date(data.date)),
#                              # expand=expansion(add=c(14,14)))+
#   expand=expansion(add=c(5,150)))+
#   coord_cartesian(ylim=c(0.1,nrow(can.var2)))+
#   # scale_color_manual(values=respcolz)+
#   scale_y_continuous(expand = expansion(add=c(0, 0)))+
#   VOCColScale
# p2
# ggsave(paste0(f.out,"Canada-variant-intervention-timeline.png"),height=1.7,width=8,units="in")

```

## Plot of total importations over time for each variant
```{r}
#import for each variant where this object was already created
alpha<-read.csv("../03_ancestral/2023-06-26_alpha_analysis/alldailyintros.csv")
alpha$var.WHO<-"Alpha"

beta<-read.csv("../03_ancestral/2023-06-26_beta_analysis/alldailyintros.csv")
beta$var.WHO<-"Beta"

gamma<-read.csv("../03_ancestral/2023-06-26_gamma_analysis/alldailyintros.csv")
gamma$var.WHO<-"Gamma"

delta<-read.csv("../03_ancestral/2023-07-06_delta_analysis/alldailyintros.csv")
delta$var.WHO<-"Delta"
#something wrong here with the singletons.. work around for now
# delta<-delta %>% filter(type == "sublineage")
# #read in an older version of singleton summary, should be the same
# delta.sing<-read.csv("../03_ancestral/2023-02-10_delta_analysis/sing.par.roll.summary.csv")
# totalsingle<-delta.sing %>% 
#   dplyr::group_by(tmrca.dt.half) %>%
#   dplyr::summarize(total.intros=sum(intros_meansum7d.mean, na.rm=T)) %>% as.data.frame()
# totalsingle$type<-"singleton"
# colnames(totalsingle)[1]<-"tmrca.dt"
# totalsingle$var.WHO<-"Delta"
# delta2<-bind_rows(totalsingle, delta)

## Add VOIs
#zeta
zeta<-read.csv("../03_ancestral/2023-07-10_zeta_analysis/alldailyintros.csv")
zeta$var.WHO<-"Zeta"

#gamma
gamma<-read.csv("../03_ancestral/2023-07-06_gamma_analysis/alldailyintros.csv")
gamma$var.WHO<-"Gamma"

#epsilon
epsilon<-read.csv("../03_ancestral/2023-07-10_epsilon_analysis/alldailyintros.csv")
epsilon$var.WHO<-"Epsilon"

#eta
eta<-read.csv("../03_ancestral/2023-07-10_eta_analysis/alldailyintros.csv")
eta$var.WHO<-"Eta"

#kappa
kappa<-read.csv("../03_ancestral/2023-07-10_kappa_analysis/alldailyintros.csv")
kappa$var.WHO<-"Kappa"

#iota
iota<-read.csv("../03_ancestral/2023-07-10_iota_analysis/alldailyintros.csv")
iota$var.WHO<-"Iota"

#mu
mu<-read.csv("../03_ancestral/2023-07-10_mu_analysis/alldailyintros.csv")
mu$var.WHO<-"Mu"


#Omicron 
ba1<-read.csv("../03_ancestral/2023-07-05_ba1_analysis/alldailyintros.csv")
ba1$var.WHO<-"BA.1"
ba11<-read.csv("../03_ancestral/2023-07-05_ba11_analysis/alldailyintros.csv")
ba11$var.WHO<-"BA.1.1"

ba2<-read.csv("../03_ancestral/2023-07-06_ba2_analysis/alldailyintros.csv")
ba2$var.WHO<-"BA.2"


# ba1.sing<-read.csv("../03_ancestral/2023-02-13_BA1_analysis/sing.par.roll.summary.csv")
# ba1.sub<-read.csv("../03_ancestral/2023-02-13_BA1_analysis/sum.par.roll.summary.csv")
# 
# ba11.sing<-read.csv("../03_ancestral/2023-02-13_BA11_analysis/sing.par.roll.summary.csv")
# ba11.sub<-read.csv("../03_ancestral/2023-02-13_BA11_analysis/sum.par.roll.summary.csv")
# 
# ba2.sing<-read.csv("../03_ancestral/2023-02-13_BA2_analysis/sing.par.roll.summary.csv")
# ba2.sub<-read.csv("../03_ancestral/2023-02-13_BA2_analysis/sum.par.roll.summary.csv")

#BA.1 and BA.1.1 together:
# BA1.sing<-bind_rows(ba1.sing,ba11.sing)
# BA1.sub<-bind_rows(ba1.sub,ba11.sub)

# function for singletons
# summ.singlez<-function(df, name){
#     #summarize singles
#   summ.single<-df%>% 
#     dplyr::group_by(tmrca.dt.half) %>%
#     dplyr::summarize(total.intros=sum(intros_meansum7d.mean, na.rm=T)) %>% as.data.frame()
#   summ.single$type<-"singleton"
#   colnames(summ.single)[1]<-"tmrca.dt"
#   summ.single$var.WHO<-name
#   return(summ.single)
# }
# # function for sublins
# summ.sublinz<-function(df, name){
#     #summarize singles
#   summ.sublin<-df%>% 
#     dplyr::group_by(tmrca.dt) %>%
#     dplyr::summarize(total.intros=sum(intros_meansum7d.mean, na.rm=T)) %>% as.data.frame()
#   summ.sublin$type<-"sublineage"
#   summ.sublin$var.WHO<-name
#   return(summ.sublin)
# }
# 
# 
# totalba.1.sing<-summ.singlez(ba1.sing, "BA.1")
# totalba.11.sing<-summ.singlez(ba11.sing, "BA.1.1")
# totalba.2.sing<-summ.singlez(ba2.sing, "BA.2")
# 
# totalba.1.sub<-summ.sublinz(ba1.sub, "BA.1")
# totalba.11.sub<-summ.sublinz(ba11.sub, "BA.1.1")
# totalba.2.sub<-summ.sublinz(ba2.sub, "BA.2")
# 
# #merge sublin and singles
# BA1.all<-bind_rows(totalba.1.sing, totalba.1.sub)
# BA11.all<-bind_rows(totalba.11.sing, totalba.11.sub)
# BA2.all<-bind_rows(totalba.2.sing, totalba.2.sub)

## AND the VOIs, cause they matter too
# eta.sing<-read.csv("../03_ancestral/     sing.par.roll.summary.csv")
# eta.sub<-read.csv("../03_ancestral/2023-02-13_BA1_analysis/sum.par.roll.summary.csv")
# 
# iota.sing<-read.csv("../03_ancestral/2023-02-13_BA11_analysis/sing.par.roll.summary.csv")
# iota.sub<-read.csv("../03_ancestral/2023-02-13_BA11_analysis/sum.par.roll.summary.csv")
# 
# zeta.sing<-read.csv("../03_ancestral/2023-02-13_BA2_analysis/sing.par.roll.summary.csv")
# zeta.sub<-read.csv("../03_ancestral/2023-02-13_BA2_analysis/sum.par.roll.summary.csv")
# 
# epsilon.sing<-read.csv("../03_ancestral/2023-02-13_BA1_analysis/sing.par.roll.summary.csv")
# epsilon.sub<-read.csv("../03_ancestral/2023-02-13_BA1_analysis/sum.par.roll.summary.csv")
# 
# mu.sing<-read.csv("../03_ancestral/2023-02-13_BA11_analysis/sing.par.roll.summary.csv")
# mu.sub<-read.csv("../03_ancestral/2023-02-13_BA11_analysis/sum.par.roll.summary.csv")
# 
# kappa.sing<-read.csv("../03_ancestral/2023-02-13_BA2_analysis/sing.par.roll.summary.csv")
# kappa.sub<-read.csv("../03_ancestral/2023-02-13_BA2_analysis/sum.par.roll.summary.csv")


# merge into one dataframe
# all.intros<-bind_rows(alpha, beta, gamma, delta, ba1, ba11, ba2)
#including VOIs
all.intros<-bind_rows(alpha, beta, gamma, delta, zeta,gamma,epsilon, eta, kappa, iota, mu, ba1, ba11, ba2)

# all.intros<-bind_rows(alpha, beta, gamma, delta2, BA1.all, BA11.all, BA2.all)

#plot together
all.intros$tmrca.dt<-as.Date(all.intros$tmrca.dt)

ylim<-c(0,600)

plot.daily.imports<-all.intros %>%
  ggplot()+
  pubThemeDate+
  theme(axis.title.y=element_text(size=rel(1)),
        legend.position="none",#"top",#c(0.3,0.66),
        legend.text=element_text(size=rel(0.9)),
        legend.title = element_text(size=rel(1.1)),
        plot.margin=margin(1,1,1,10,unit="pt"))+
        # axis.text.x=element_blank())+
  geom_bar(aes(x=tmrca.dt,y=total.intros,fill=var.WHO),stat="identity",width=1)+
  labs(x=NULL, y="Daily variant introductions into Canada")+
  scale_y_continuous(expand=expansion(c(0,0.02)))+
  VarFillScaleRed2+
  VarColScaleRed2+
  scaleDateBusy+
  guides(fill=guide_legend(title.position = "left",keywidth = 1,keyheight = 1, 
                           title="Variant",ncol=4))+
  #ADD INTERVENTIONS
  #description and lines to axis
  geom_segment(data=can.var2, aes(y=position2+ylim[2],yend=position2+ylim[2], 
                              x=Start.date,xend=End.date,color=var.WHO),
               size=3,alpha=1 )+
  geom_text(data=can.var2,aes(x=Start.date,label=Description,y=position2+50+ylim[2],
                              hjust=hjusts), fontface="bold", size=3)
#plot dat
plot.daily.imports
  
```

## Make a version that is only the lower axis range (pre-omi)
```{r}

ylim2<-c(0,150)

plot.daily.imports.less<-all.intros %>%
  ggplot()+
  pubThemeDate+
  theme(axis.title.y=element_text(size=rel(1)),
        legend.position="none",#"top",#c(0.3,0.66),
        legend.text=element_text(size=rel(0.9)),
        legend.title = element_text(size=rel(1.1)),
        plot.margin=margin(1,1,1,10,unit="pt"))+
        # axis.text.x=element_blank())+
  geom_bar(aes(x=tmrca.dt,y=total.intros,fill=var.WHO),stat="identity",width=1)+
  labs(x=NULL, y="Daily variant introductions into Canada")+
  scale_y_continuous(expand=expansion(c(0,0.02)))+
  VarFillScaleRed2+
  VarColScaleRed2+
  scaleDateBusy+
  guides(fill=guide_legend(title.position = "left",keywidth = 1,keyheight = 1, 
                           title="Variant",ncol=4))+
  #ADD INTERVENTIONS
  #description and lines to axis
  geom_segment(data=can.var2, aes(y=position2+ylim2[2],yend=position2+ylim2[2], 
                              x=Start.date,xend=End.date,color=var.WHO),
               size=3,alpha=1 )+
  geom_text(data=can.var2,aes(x=Start.date,label=Description,y=position2+50+ylim2[2],
                              hjust=hjusts), fontface="bold", size=3)+
  coord_cartesian(ylim=ylim2)
  #plot dat
plot.daily.imports.less
ggsave(paste0(f.out,"/Figure1B_zoom.png"),height=4,width=4)

```


## Read in the png of the cartoon phylo and grob everything together with A-C
```{r}
p.phylo<-cowplot::ggdraw() + cowplot::draw_image("Figure1_plot/cartoon-phylo/cartoon-phylo-BA.png", scale = 1)

plotrow<-plot_grid(plot.daily.ave.voc.cases.fplus, #version with waves
          plot.daily.imports,
          labels=LETTERS[1:2], nrow=2, rel_heights=c(0.8,1),
          align="v",label_x=-0.01) #extra room for axis

#now add the cartoon phylo  
plot_grid(plotrow, p.phylo, labels=c(NA, "C"),
          label_x=0.05, label_y=0.85,
          rel_widths = c(1,0.6),align="h", axis="topleft",
          rel_heights=c(1,1))


ggsave(paste0(f.out,"/Figure1_composite.png"),height=6,width=8.5)

```

