---
title: "Untitled"
author: "Angela"
date: '2023-01-03'
output: html_document
---

Objectives
* compare the root-to-tip regression of 
1) variants, 
2) subsampling temporal vs uniform, 
3) bootstraps

* extract/re-calculate original slopes pre-exclusion and post-exclusion 
* make a single slope across boots to compare variants' rtt
* compare the inferred strict and relaxed clock rates by variant
* compare the inferred clock rate from strict rtt and relaxed model for 1, 2, 3
* sampling proportion matters... go back to 2020 analysis to compare different % inclusion of canadian seqs

```{r}
# setup libraries
library(stringr)
library(ape)
library(treeio)
library(adephylo)
library(tidyverse)
library(gtools)
library(lubridate)
library(scales)
library(reshape2)

#set up outputs
f.out<-"compare_rtt"
if(!dir.exists(f.out)){dir.create(f.out)}
```

## plotting extras
```{r}
data.date<-"2022-03-22" 
start.date<-"2020-11-01"

pubThemeDate<-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    panel.background=element_rect("grey95"), axis.line = element_line(colour = "black"),
                    text=element_text(size=10,face="bold"),
                    legend.key.size = unit(0.4,"line"),
                    legend.text=element_text(size=8),
                    axis.text.x=element_text(angle = 45,hjust=1,size=rel(1)))

xdates<-ymd(c("2020-12-01","2021-03-01","2021-06-01","2021-09-01","2021-12-01","2022-03-01"))
scaleDateBusy<-scale_x_date(breaks=xdates, date_labels = "%b %Y",
                            limits=as.Date(c(start.date,"2022-03-22")),
                            expand=c(0,0))

variant.colors.in<-"../../01_subSample/CleanMeta_PreSubsamp/variant_colors.tsv"
variant.df.in<-paste0("../../01_subSample/Results_PreSubsamp/variant.df.can.csv")

## read in the variant df to use for inputs
var.df<-read.csv(variant.df.in)
var.df<-var.df[with(var.df, order(var.df$first.can.date, decreasing = F)),]
#add to var.df
var.omi<-var.df[which(var.df$var.WHO=="Omicron"),]
var.omi<-rbind(var.omi,var.omi,var.omi,var.omi) #one for each sep lineage considered
omis<-c("BA.1","BA.1.1","BA.2", "BA.3")
var.omi$var.WHO<-omis
var.omi$var.lin<-omis
var.omi$first.glob.date<-c("2021-09-02","2021-09-09","2021-09-21","2021-11-23")
#join var.omi onto var.df 
var.df<-rbind(var.df,var.omi)
#remove 'omicron' full group
var.df<-var.df[-which(var.df$var.WHO=="Omicron"),]

#take the names
var.ord<-var.df$var.WHO
var.ord<-var.ord[-which(var.ord %in% c("Theta","Lambda","GH/490R"))]

#read in variant colors
varPalette<-read.table(variant.colors.in,sep="\t")

## make color scheme
var.colz<-as.character(varPalette$vars.ch)
varPalette.ch<-row.names(varPalette)
names(varPalette.ch)<-var.colz

### add separate groups for each omicron group
omis<-c("BA.1","BA.1.1","BA.2")
# https://www.colorhexa.com/ec9a91
omiPalette.ch<-c("#e47266","#e8867b","#f0aea7") 
names(omiPalette.ch)<-omis
morevarPal<-c(varPalette.ch[match(var.ord,names(varPalette.ch))],
              omiPalette.ch)
morevarPal<-morevarPal[-which(names(morevarPal)=="Omicron")]

var.ord.omi<-var.ord
var.ord.omi<-var.ord.omi[-which(var.ord.omi=="Omicron")]
var.ord.omi<-c(var.ord.omi,omis)
VarColScale<-scale_colour_manual(name = "Variant",values = morevarPal,na.value=NULL)
VarFillScale<-scale_fill_manual(name = "Variant",values = morevarPal,na.value=NULL)

```

# FOR EACH VARIANT
```{r}
focalvar<- "other"
Focalvar<- "Other"
treebase<-"../202210_analysis/ft_other_root_res" #locaiton of residuals/rtt dist
resbase<-"../202210_analysis/ft_other_root" #locaiton of residuals/rtt dist
metabase<-"../../01_subSample/bootsamples_meta_202210" 

# focalvar<- "delta"
# Focalvar<- "Delta"
# treebase<-"../202208_analysis/ft_alphadelta_root_res" #locaiton of residuals/rtt dist
# resbase<-"../202208_analysis/ft_alphadelta_root" #locaiton of residuals/rtt dist
# metabase<-"../../01_subSample/bootsamples_meta_202208" #might not need
```

## Run for each
```{r}
#setup treefiles
trees.in<-list.files(treebase,pattern = "res.tre",full.names = T) %>% mixedsort ()
if(length(trees.in)<1 | all(!file.exists(trees.in))){
    print("trees not found");break
}

#restrict to focus VOC
trees.in<-str_subset(trees.in,paste0("/",focalvar,"_"))

#number of bootstraps/trees
b<-length(trees.in)

#list residual data
tempest.in<-str_replace_all(trees.in, "_res.tre",".tsv")
tempest.in<-str_replace_all(tempest.in,treebase, resbase)
if(length(tempest.in)<1 | all(!file.exists(tempest.in))){
    print("tempest tsv not found");break
}

#metadata
meta.in<-str_replace_all(trees.in, treebase,metabase)
meta.in<-str_replace_all(meta.in, "_mask_root_res.tre",".csv")
if(length(meta.in)<1 | all(!file.exists(meta.in))){
  print("meta not found");break
}

## Read in rtt dist and dates for clean subs scaled trees

# Read in rooted/excl trees (output from Remove outliers), residual (r-t-t dist) and dates,
## Read in the trees into a list
trees<-replicate(n=b, vector)

for (i in 1:b){
  trees[[i]]<-read.tree(file=trees.in[i])
}

## Read in tempest data with residuals here
resid<-replicate(n=b,vector())
for (i in 1:b){
  resid[[i]]<-read.table(file=tempest.in[i] ,sep="\t",header=T)
  #boot name
  boot.nm<-str_replace_all(last(unlist(strsplit(tempest.in[i],"/"))),"_mask_root.tsv","")
  resid[[i]]$boot<-boot.nm
}

# exclude resid rows if not in the tree
for (i in 1:b){
  nope<-which(!resid[[i]]$tip %in% trees[[i]]$tip.label)
  if(length(nope)>0){
    resid[[i]]<-resid[[i]][-nope,]
  }
}

# read in the meta
meta<-replicate(n=b,vector())
for (i in 1:b){
  meta[[i]]<-read.table(file=meta.in[i] ,sep=",",header=T)
  meta[[i]]$boot<-i
  meta[[i]]$tip<-meta[[i]]$new.names
  #calculate meta decdates
  meta[[i]]$decdate<-decimal_date(as.Date(meta[[i]]$date))
  #change this so we can use later
  meta[[i]]$incomp[meta[[i]]$country!="Canada"]<-0
  #trim the object down
  meta[[i]]<-meta[[i]][,c("tip","var.WHO","Lineage","decdate","incomp","date")]
  colnames(meta[[i]])[ncol(meta[[i]])]<-"fulldate"
}

#Link to resid by tiplabel
for (i in 1:b){
  resid[[i]]<-left_join(resid[[i]], meta[[i]],by="tip")
  #replace the resid date with the meta decdates
  resid[[i]]$date<-resid[[i]]$decdate
  resid[[i]]<-resid[[i]][,-which(colnames(resid[[i]])=="decdate")]
}

#restrict to variants
if(focalvar != "other"){ #as long as not analyzing Other specifically:
  for (i in 1:b){
    if(any(is.na(resid[[i]]$var.WHO))){ #NA
      resid[[i]]<-resid[[i]][-which(is.na(resid[[i]]$var.WHO)),]
    }
    if(any(resid[[i]]$var.WHO=="Other")){ 
      resid[[i]]<-resid[[i]][-which(resid[[i]]$var.WHO=="Other"),]
    }
  }
}

## Make a plot of rtt regression for each variant
##make a linear fit for each boot
summary.table<-data.frame(focalvar=rep(Focalvar,10), 
                          boot=1:b, 
                          n=NA,
                          slope=NA,
                          intercept=NA,
                          eqn=NA,
                          adjR2=NA)
for (i in 1:b){
  m<-lm(resid[[i]]$distance ~ resid[[i]]$date)
  # plot(resid[[i]]$distance ~ resid[[i]]$date)
  #calculate the residuals and replace in the df
  resid[[i]]$residual<-residuals(m)
  #extract slope
  lm<-summary(m)
  slope.n<-coef(lm)[[2]] #numeric
  slope<-formatC(slope.n,format="e", digits=2)

  # add to summary table
  summary.table$n[i]<-nrow(resid[[i]])
  summary.table$slope[i]<-slope.n
  summary.table$intercept[i]<-coef(lm)[[1]]
  summary.table$eqn[i]<-paste("y = ",slope," x + ",signif(int,digits=3),sep="")
  summary.table$adjR2[i]<-signif(lm$adj.r.squared,digits = 3)
}

if(focalvar=="delta"){
  #delta has 3 bad boots: 4, 7, 9
  summary.table<-summary.table[-c(4,7,9),]
}

#downsample for plot
restip<-sample(resid[[1]]$tip,size=nrow(resid[[1]])*0.5, replace = F)
ressamp<-resid[[1]][match(restip,resid[[1]]$tip),]
```  

## save as separate objects
```{r}
summary.table.other<-summary.table
ressamp.other<-ressamp
```

# join variants together
```{r}
summary.table.all<-rbind(summary.table.ba2,summary.table.ba1,summary.table.ba11,
                         summary.table.alpha, summary.table.delta, summary.table.beta,
                         summary.table.gamma,summary.table.iota,summary.table.eta,
                         summary.table.epsilon,summary.table.kappa,summary.table.mu,
                         summary.table.zeta, summary.table.other
                         )
ressamp.all<-rbind(ressamp.ba2,ressamp.ba1,ressamp.ba11,
                   ressamp.alpha,ressamp.delta,ressamp.beta,
                   ressamp.gamma,ressamp.iota,ressamp.eta,
                   ressamp.epsilon,ressamp.kappa,ressamp.mu,
                   ressamp.zeta, ressamp.other
                   )
```

## PLOT STRICT CLOCK RATES violin and RTT across vars
```{r}
summary.table.all$focalvar<-factor(summary.table.all$focalvar,var.ord.omi)
#Plot a violin plot of strict molecular clock rate
v.p<-ggplot(summary.table.all)+
  geom_point(aes(x=focalvar, y=slope),size=rel(0.7))+    
  geom_violin(aes(x=focalvar, y=slope,fill=focalvar),alpha=0.7,width=1,scale="width")+
  labs(x=NULL,y="Strict clock rate [s/s/y]")+
  scale_y_continuous(breaks=seq(0,0.0015,0.0002),labels = format(seq(0,0.0015,0.0002),scientific = T))+
  guides(fill=guide_legend(title="Variant"))+
  pubThemeDate+
  theme(legend.position = "none")+
  VarFillScale
v.p
ggsave(paste0(f.out,"/violin_strict.png"),height=3,width=4,units="in")

#downsample a random bootstrap to 50%
ressamp.all$var.WHO<-factor(ressamp.all$var.WHO,levels=var.ord.omi)
ggplot(ressamp.all,aes(x=date,y=distance, group=var.WHO, color=var.WHO))+
  geom_point(alpha=0.4)+
  labs(x="Year",y="Root-to-tip distance",color="Variant")+
  geom_smooth(method="lm", se = F,size = 2, color = "black")+
  geom_smooth(method="lm", se = TRUE)+
  pubThemeDate+
  theme(legend.position = c(0.2,0.7))+
  VarColScale
ggsave(paste0(f.out,"/rtt_all.png"),height=4,width=5,units="in")

```

## Test if slopes sig'ly different in rtt analysis?
```{r}
#kruskal: any diff?
kruskal.test(data=summary.table.all, slope ~ focalvar) #signif
#which pairs of groups are different: wilcox
df_wil<-pairwise.wilcox.test(summary.table.all$slope, summary.table.all$focalvar,
                 p.adjust.method = "BH")
df_p<-df_wil$p.value

## make a heat map of this instead of trying to overlay
## group into signif categories
mat<-df_p
mat2<-mat
for(i in 1:nrow(mat)){
  for (j in 1:ncol(mat)){
    v<-mat2[i,j]
    if(is.na(v)){next}
    if (v<1e-4){mat[i,j]<-"<1e-4";next}
    if (v>=1e-4 & v<1e-3){mat[i,j]<-"1e-4 - 1e-3";next}
    if (v>=1e-3 & v<1e-2){mat[i,j]<-"1e-3 - 1e-2";next}
    if (v>=1e-2 & v<1e-1){mat[i,j]<-"1e-2 - 1e-1";next}
    if (v>=1e-1){mat[i,j]<-">1e-1"}
  }
}
#colorss
sigs<-c("<1e-4", "1e-4 - 1e-3", "1e-3 - 1e-2", "1e-2 - 1e-1", ">1e-1")
colz<-viridis_pal(direction = -1)(length(sigs))
names(colz)<-sigs

#melt long
mat2<-mat %>% reshape2::melt()
colnames(mat2)<-c("Var2","Var1","value")
mat3<-rbind(mat2)

#remove same
rem<-c()
for (i in 1:nrow(mat3)){
  if(mat3$Var1[i]==mat3$Var2[i]){rem<-c(rem,i)}
}
length(rem)
mat4<-mat3[-rem,]

#order Var1 and Var2
mat4$Var2<-factor(mat4$Var2, levels = var.ord)
mat4$Var1<-factor(mat4$Var1, levels = var.ord)
mat4<-mat4[-which(is.na(mat4$value)),]
mat4$value<-factor(mat4$value,levels=sigs)

P.heat<- mat4 %>% ggplot()+
  geom_tile(aes(x=Var1,y=Var2,fill=factor(value)))+
  labs(x=NULL,y=NULL,fill="Pairwise Wilcox\nAdjusted p value")+
  scale_fill_manual(values=colz)+
  pubThemeDate+
  theme(legend.position="right",
           legend.text=element_text(size=10))
P.heat
ggsave(paste0(f.out,"/StrictClockHeatmap.png"),height=4,width=5,units="in")

```

## REad in the relaxed clocks - need to be parsed
```{r}
rel<-read.csv("relaxedRates.csv")
# colnames(rel)

#parse
rel$mean<-as.numeric(NA)
rel$upper<-as.numeric(NA)
rel$lower<-as.numeric(NA)
for (i in 1:nrow(rel)){
  par<-unlist(str_split(rel$string[i],","))
  par.rt<-unlist(str_split(trimws(par[1]),"\\ "))
  rel$mean[i]<-par.rt[2] 
  rel$lower[i]<-str_replace_all(par.rt[3],"\\[|;","")
  rel$upper[i]<-str_replace_all(par.rt[4],"\\]","")
}
rel$mean<-as.numeric(rel$mean)
rel$lower<-as.numeric(rel$lower)
rel$upper<-as.numeric(rel$upper)

rel$Variant<-str_to_title(rel$variant)
rel$Variant<-str_replace_all(rel$Variant,"Ba","BA")
rel$Variant<-factor(rel$Variant,var.ord.omi)
#Plot a violin plot of strict molecular clock rate
v.rel<-ggplot(data=rel,aes(x=Variant, y=mean,fill=Variant))+
  geom_point(size=rel(0.7),position=position_dodge2(width=0.5),alpha=0.5)+    
  geom_linerange(aes(ymin=lower,ymax=upper),
                position=position_dodge2(width=0.5,padding=0.5),alpha=0.5)+
  geom_violin(alpha=0.7,width=1,scale="width")+
  labs(x=NULL,y="Mean relaxed clock rate [s/s/y]")+
  scale_y_continuous(breaks=seq(0,0.001,0.0001),labels = format(seq(0,0.001,0.0001),scientific = T))+
  guides(fill=guide_legend(title="Variant"))+
  pubThemeDate+
  theme(legend.position = "none")+
  VarFillScale+
  VarColScale
v.rel
ggsave(paste0(f.out,"/violin_relaxed.png"),height=3,width=4,units="in")

#also compare relaxed to strict
```

## stats for relaxed clocks
```{r}
#kruskal: any diff?
kruskal.test(data=rel, mean ~ Variant) #signif
#which pairs of groups are different: wilcox
df_wil<-pairwise.wilcox.test(rel$mean, rel$Variant,
                 p.adjust.method = "BH")
df_p<-df_wil$p.value

## make a heat map of this instead of trying to overlay
## group into signif categories
mat<-df_p
mat2<-mat
for(i in 1:nrow(mat)){
  for (j in 1:ncol(mat)){
    v<-mat2[i,j]
    if(is.na(v)){next}
    if (v<1e-4){mat[i,j]<-"<1e-4";next}
    if (v>=1e-4 & v<1e-3){mat[i,j]<-"1e-4 - 1e-3";next}
    if (v>=1e-3 & v<1e-2){mat[i,j]<-"1e-3 - 1e-2";next}
    if (v>=1e-2 & v<1e-1){mat[i,j]<-"1e-2 - 1e-1";next}
    if (v>=1e-1){mat[i,j]<-">1e-1"}
  }
}
#colorss
sigs<-c("<1e-4", "1e-4 - 1e-3", "1e-3 - 1e-2", "1e-2 - 1e-1", ">1e-1")
colz<-viridis_pal(direction = -1)(length(sigs))
names(colz)<-sigs

#melt long
mat2<-mat %>% reshape2::melt()
colnames(mat2)<-c("Var2","Var1","value")
mat3<-rbind(mat2)

#remove same
rem<-c()
for (i in 1:nrow(mat3)){
  if(mat3$Var1[i]==mat3$Var2[i]){rem<-c(rem,i)}
}
length(rem)
mat4<-mat3[-rem,]

#order Var1 and Var2
mat4$Var2<-factor(mat4$Var2, levels = var.ord)
mat4$Var1<-factor(mat4$Var1, levels = var.ord)
mat4<-mat4[-which(is.na(mat4$value)),]
mat4$value<-factor(mat4$value,levels=sigs)

P.heatr<- mat4 %>% ggplot()+
  geom_tile(aes(x=Var1,y=Var2,fill=factor(value)))+
  labs(x=NULL,y=NULL,fill="Pairwise Wilcox\nAdjusted p value")+
  scale_fill_manual(values=colz)+
  pubThemeDate+
  theme(legend.position="right",
           legend.text=element_text(size=10))
P.heatr
ggsave(paste0(f.out,"/RelaxedClockHeatmap.png"),height=4,width=5,units="in")

```

# Comapre the different subsamp strategies taken in SC2 2020 paper
```{r}
strict20<-read.csv("../../../20210301_phylogeo/02_trees/SummaryTables/BootsSlopes.csv")

#instead of Variants, compare across strategies (25 vs 50 75 100%)
head(strict20) #slopefinal
strict20$slope.final<-as.numeric(strict20$slope.final)
colnames(strict20)
```

## PLOT STRICT rates 2020 violin across strategies
```{r}
strat.ord<-c("25%","50%","75%","100%")
strict20$Strategy<-factor(strict20$Strategy,strat.ord)
#Plot a violin plot of strict molecular clock rate
v.s20<-ggplot(strict20)+
  geom_point(aes(x=Strategy, y=slope.initial),size=rel(0.7))+    
  geom_violin(aes(x=Strategy, y=slope.initial,fill=Strategy),alpha=0.7,width=1,scale="width")+
  labs(x="% Canadian seqs included",y="Strict clock rate [s/s/y]")+
  scale_y_continuous(breaks=seq(0,0.0015,0.0002),labels = format(seq(0,0.0015,0.0002),scientific = T))+
  pubThemeDate+
  theme(legend.position = "none")
v.s20
ggsave(paste0(f.out,"/2020_strat_violin_strict.png"),height=3,width=4,units="in")

## SIGNIFI?
kruskal.test(data=strict20, slope.initial ~ Strategy) #signif
#which pairs of groups are different: wilcox
df_wil<-pairwise.wilcox.test(strict20$slope.initial, strict20$Strategy,
                 p.adjust.method = "BH")
df_p<-df_wil$p.value

## make a heat map of this instead of trying to overlay
## group into signif categories
mat<-df_p
mat2<-mat
for(i in 1:nrow(mat)){
  for (j in 1:ncol(mat)){
    v<-mat2[i,j]
    if(is.na(v)){next}
    if (v<1e-4){mat[i,j]<-"<1e-4";next}
    if (v>=1e-4 & v<1e-3){mat[i,j]<-"1e-4 - 1e-3";next}
    if (v>=1e-3 & v<1e-2){mat[i,j]<-"1e-3 - 1e-2";next}
    if (v>=1e-2 & v<1e-1){mat[i,j]<-"1e-2 - 1e-1";next}
    if (v>=1e-1){mat[i,j]<-">1e-1"}
  }
}
#colorss
sigs<-c("<1e-4", "1e-4 - 1e-3", "1e-3 - 1e-2", "1e-2 - 1e-1", ">1e-1")
colz<-viridis_pal(direction = -1)(length(sigs))
names(colz)<-sigs

#melt long
mat2<-mat %>% reshape2::melt()
colnames(mat2)<-c("Var2","Var1","value")
mat3<-rbind(mat2)

#remove same
rem<-c()
for (i in 1:nrow(mat3)){
  if(mat3$Var1[i]==mat3$Var2[i]){rem<-c(rem,i)}
}
length(rem)
mat4<-mat3[-rem,]

#order Var1 and Var2
mat4$Var2<-factor(mat4$Var2, levels = strat.ord)
mat4$Var1<-factor(mat4$Var1, levels = strat.ord)
mat4<-mat4[-which(is.na(mat4$value)),]
mat4$value<-factor(mat4$value,levels=sigs)

P.heat20<- mat4 %>% ggplot()+
  geom_tile(aes(x=Var1,y=Var2,fill=factor(value)))+
  labs(x=NULL,y=NULL,fill="Pairwise Wilcox\nAdjusted p value")+
  scale_fill_manual(values=colz)+
  pubThemeDate+
  theme(legend.position="right",
           legend.text=element_text(size=10))
P.heat20
ggsave(paste0(f.out,"/2020_StrictClockHeatmap.png"),height=2,width=3,units="in")

```

## PLOT Relaxed rates 2020 violin across strategies
```{r}
rel20<-read.csv("../../../20210301_phylogeo/02_trees/SummaryTables/tmrca_log_parsed.csv")
rel20$Strategy<-factor(rel20$Strategy,strat.ord)
rel20<-rel20[-which(is.na(rel20$Strategy)),]

rel20$Mean<-as.numeric(rel20$Mean) #some NAs
rel20$lower.95..CI<-as.numeric(rel20$lower.95..CI)
rel20$upper.95..CI<-as.numeric(rel20$upper.95..CI)

#Plot a violin plot of Relaxed molecular clock rate
v.rel20<-ggplot(data=rel20,aes(x=Strategy, y=Mean,fill=Strategy))+
  geom_point(size=rel(0.7),position=position_dodge2(width=0.5),alpha=0.5)+    
  geom_linerange(aes(ymin=lower.95..CI,ymax=upper.95..CI),
                position=position_dodge2(width=0.5,padding=0.5),alpha=0.5)+
  geom_violin(alpha=0.7,width=1,scale="width")+
  labs(x=NULL,y="Mean relaxed clock rate [s/s/y]")+
  scale_y_continuous(breaks=seq(0,0.001,0.0001),labels = format(seq(0,0.001,0.0001),scientific = T))+
  labs(x="% Canadian seqs included",y="Relaxed clock rate [s/s/y]")+
  pubThemeDate+
  theme(legend.position = "none")
v.rel20
ggsave(paste0(f.out,"/2020_strat_violin_Relaxed.png"),height=3,width=4,units="in")

## SIGNIFI?
kruskal.test(data=rel20, Mean ~ Strategy) #signif p-value = 0.01346
#which pairs of groups are different: wilcox
df_wil<-pairwise.wilcox.test(rel20$Mean, rel20$Strategy,
                 p.adjust.method = "BH")
df_p<-df_wil$p.value

## make a heat map of this instead of trying to overlay
## group into signif categories
mat<-df_p
mat2<-mat
for(i in 1:nrow(mat)){
  for (j in 1:ncol(mat)){
    v<-mat2[i,j]
    if(is.na(v)){next}
    if (v<1e-4){mat[i,j]<-"<1e-4";next}
    if (v>=1e-4 & v<1e-3){mat[i,j]<-"1e-4 - 1e-3";next}
    if (v>=1e-3 & v<1e-2){mat[i,j]<-"1e-3 - 1e-2";next}
    if (v>=1e-2 & v<1e-1){mat[i,j]<-"1e-2 - 1e-1";next}
    if (v>=1e-1){mat[i,j]<-">1e-1"}
  }
}
#colorss
sigs<-c("<1e-4", "1e-4 - 1e-3", "1e-3 - 1e-2", "1e-2 - 1e-1", ">1e-1")
colz<-viridis_pal(direction = -1)(length(sigs))
names(colz)<-sigs

#melt long
mat2<-mat %>% reshape2::melt()
colnames(mat2)<-c("Var2","Var1","value")
mat3<-rbind(mat2)

#remove same
rem<-c()
for (i in 1:nrow(mat3)){
  if(mat3$Var1[i]==mat3$Var2[i]){rem<-c(rem,i)}
}
length(rem)
mat4<-mat3[-rem,]

#order Var1 and Var2
mat4$Var2<-factor(mat4$Var2, levels = strat.ord)
mat4$Var1<-factor(mat4$Var1, levels = strat.ord)
mat4<-mat4[-which(is.na(mat4$value)),]
mat4$value<-factor(mat4$value,levels=sigs)

P.heat20<- mat4 %>% ggplot()+
  geom_tile(aes(x=Var1,y=Var2,fill=factor(value)))+
  labs(x=NULL,y=NULL,fill="Pairwise Wilcox\nAdjusted p value")+
  scale_fill_manual(values=colz)+
  pubThemeDate+
  theme(legend.position="right",
           legend.text=element_text(size=10))
P.heat20
ggsave(paste0(f.out,"/2020_RelaxedClockHeatmap.png"),height=2,width=3,units="in")

```

# UNIFORM temporal
```{r}
### for other, have this comaprison
other<-summary.table.other
other<-other[,c(2,4)]
other$temp<-"Case proportional"

table2<-strict20[strict20$Strategy=="50%",]
table2<-table2[,c("Bootstrap","slope.final")]
table2$temp<-"Uniform"
colnames(table2)<-colnames(other)

other.strict<-bind_rows(other,table2)
other.strict$clock<-"strict"

#relaxed temp
other.rel<-rel[rel$Variant=="Other",]
other.rel<-other.rel[,c("boot","mean")]
other.rel$temp<-"Case proportional"
table3<-rel20[rel20$Strategy=="50%",]
table3<-table3[,c("Bootstrap","Mean")]
table3$temp<-"Uniform"
colnames(table3)<-colnames(other.rel)
other.rel2<-bind_rows(other.rel,table3)
other.rel2$clock<-"relaxed"

colnames(other.strict)<-colnames(other.rel2)
all.other<-bind_rows(other.strict,other.rel2)

other.p<-ggplot(all.other,aes(x=clock,fill=temp,y=mean))+
  geom_point(size=rel(0.6),position = position_dodge2(width=0.7),alpha=0.5)+    
  geom_violin(alpha=0.7,width=0.7,scale="width")+
  labs(x=NULL,y="Clock rate [s/s/y]")+
  scale_y_continuous(breaks=seq(0,0.0015,0.0002),labels = format(seq(0,0.0015,0.0002),scientific = T))+
  guides(fill=guide_legend(title="Temporal\nsampling\nstrategy"))+
  pubThemeDate+
  theme(legend.position = c(0.25,0.8))+
  annotate(geom="text",x = 1,y=3.6e-4, label="p=1.5e-4",size=rel(2))+
  annotate(geom="segment",x=0.75,xend=1.25,y=3.4e-4,yend=3.4e-4)+
  annotate(geom="text",x = 2,y=8.6e-4, label="p=0.016",size=rel(2))+
  annotate(geom="segment",x=1.75,xend=2.25,y=8.4e-4,yend=8.4e-4)

  other.p
ggsave(paste0(f.out,"/violin_other.png"),height=3,width=4,units="in")

all.other.2<-all.other[all.other$clock=="relaxed",]
all.other.1<-all.other[all.other$clock=="strict",]

kruskal.test(data=all.other.1, mean ~ temp) #signif  p-value = 0.01553
kruskal.test(data=all.other.2, mean ~ temp) #signif  p-value = 0.0001571
#which pairs of groups are different: wilcox
df_wil<-pairwise.wilcox.test(all.other.1$mean, all.other.1$temp,
                 p.adjust.method = "BH") #0.017
df_wil<-pairwise.wilcox.test(all.other.2$mean, all.other.2$temp,
                 p.adjust.method = "BH") #1.1e-05 
```

